---
title: "categoriseSubsetCVDDeaths"
output: html_document
---

  Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and import data                                                  
             (B) Derive causes using ICD codes
             (C) Count number of deaths and save file


## Load the relevant libraries and import data   

```{r}
library("dplyr")
library("tidyr")
library("data.table")

load("R:/working/preparing data/DeathData/subsetcvd/icd10/deathsbyCauseICD10.RDATA")
load("R:/working/preparing data/DeathData/subsetcvd/icd9/deathsbyCausesICD9_final.RDATA")
deaths<-rbind(deathsbyCauses9final,deathsbyCauses10final)

```


## Derive causes using ICD codes

```{r}
deaths$cause<-trimws(deaths$cause)

deaths_ischemic<- deaths %>% filter(cause=="ischemic")
deaths_stroke<- deaths %>% filter(cause=="stroke")


deaths<-deaths_ischemic

```


## Count number of deaths and save file

```{r}
#Count number of deaths
mortcounts<-count(deaths, year, CT, sex, age)
colnames(mortcounts)[colnames(mortcounts)=="n"]<-"death"

#sort data by sequentially of CT
deaths<-mortcounts[order(mortcounts$CT),]


save(deaths, file="R:/working/preparing data/DeathData/causes/subset/deaths_subset.rdata")

```

