---
title: "categorise_ICD10"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Assess patterns of ICD codes and categorise death files back into GBD cause groups                                              
             (C) Create ICD9 list                                                                      
             (D) Load all the datasets needed to restrict the death data
             (E) Reassign CT of mortality data based on number of years spent in residency

## Load libraries, import data, and define attributes   

```{r}
library(icd)
library(tidyverse)
load("R:/working/preparing data/DeathData/causes/deaths_icd10.Rdata")
dir<-"R:/working/preparing data/DeathData/subsetcvd/icd10/"
setwd(dir)
causefiles=(list.files(pattern="*.csv"))
load("R:/working/preparing data/DeathData/cvdcausesicd10.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)
```

## Assess patterns of ICD codes and categorise death files back into GBD cause groups   

```{r}
#trial with just three characters
trial<-sapply(gbdCausesicd10, function(x) ifelse(nchar(x)>3, substr(x,1,nchar(x)-1), x))
deaths_icd10$code<-ifelse(nchar(deaths_icd10$ICD10)>3, substr(deaths_icd10$ICD10,1,nchar(deaths_icd10$ICD10)-1), deaths_icd10$ICD10)

deaths_icd10<-subset(deaths_icd10, select = -c(icdversn, causeidgarb))

deaths_icd10$newcause<-0
for (y in causenames) {
  deaths_icd10$newcause[which(deaths_icd10$code %in% trial[[y]])] <- y
}

deaths_icd10$cause<-0
deaths_icd10$cause<-ifelse(deaths_icd10$causeid==0, deaths_icd10$newcause, deaths_icd10$causeid)

deaths_icd10 %>% filter(cause=="stroke" | cause=="ischemic") %>% tally() #4114 counts

#Next trial with just two characters
trial2<-sapply(gbdCausesicd10, function(x) ifelse(nchar(x)>2, substr(x,1,nchar(x)-2), x))
trial2<-sapply(trial2, function(x) ifelse(nchar(x)<2,NA, x))
deaths_icd10$code2<-ifelse(nchar(deaths_icd10$ICD10)>3, substr(deaths_icd10$ICD10,1,nchar(deaths_icd10$ICD10)-2), deaths_icd10$ICD10)
deaths_icd10$code2<-ifelse(nchar(deaths_icd10$code2)>2, substr(deaths_icd10$code2,1,nchar(deaths_icd10$code2)-1), deaths_icd10$code2)

deaths_icd10$newcause2<-0
for (y in causenames) {
  deaths_icd10$newcause2[which(deaths_icd10$code2 %in% trial2[[y]])] <- y
}
deaths_icd10$causes<-0
deaths_icd10$causes<-ifelse(deaths_icd10$cause==0, deaths_icd10$newcause2, deaths_icd10$cause)

deaths_icd10 %>% filter(cause=="stroke" | cause=="ischemic") %>% tally() #4114

save(deaths_icd10, file="R:/working/preparing data/DeathData/subsetcvd/icd10/deaths_icd10.RDATA")

#Tally up
deathsbyCause10<-deaths_icd10 %>%
  group_by(newcause) %>%
  tally()

deathsbyCauseFinal<-deaths_icd10 %>%
  group_by(causes) %>%
  tally()

#Isolate causes with no causes identified
noCauses<-subset(deaths_icd10, causes==0)
noCauses<-select(noCauses,1:5)
colnames(noCauses)[5]<-"code"
noCauses<-left_join(noCauses, icd10cm2019, by="code")

noCauseTally<-noCauses %>%
  group_by(major) %>%
  tally()

deathsbyCauses10final<-select(deaths_icd10, year, CT, sex, age, newcause2) %>% rename(cause=newcause2)
deathsbyCauses10final<-deathsbyCauses10final %>% filter(cause=="stroke" | cause=="ischemic")  #52233

save(deathsbyCauses10final, file="R:/working/preparing data/DeathData/subsetcvd/icd10/deathsbyCauseICD10.RDATA")
```

