---
title: "PrepareCauses_ICD10"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Prepare ICD codes for 21 GBD causes     

## Load libraries, import data, and define attributes 

```{r}
library(stringr)
library(abind)
library(data.table)
library(dplyr)
library(stringi)
dir<-"R:/working/preparing data/DeathData/subsetcvd/icd10/"
setwd(dir)
```

## Prepare ICD codes for 21 GBD causes

```{r}
causefiles=(list.files(pattern="*.csv"))
causefiles=setNames(causefiles, substr(list.files(pattern="*.csv"),1,nchar(causefiles)-4))
allCauses<-lapply(causefiles,read.table,header=F,sep=",", stringsAsFactors=F)
transposer<-function(df){
  z<-t(df)
}
allCauses<-lapply(allCauses, transposer)

##FIND OUT how to apply this function to the list
causes=function(x,output){
  V1 = x[1]
  V2 = gsub("[A-Z]","", V1)
  V3 = substring(V1,1,1)
  V4 = word(V2,1,sep="\\-")
  V5 = sub("^.+-","", V2)
  seq<-as.data.frame((t(seq(V4,V5, by=0.1))))
  final = sub("^",V3, seq)
  final = gsub("\\.","",final)
}

allCauses2<-do.call(rbind.data.frame, allCauses)
allCauses3<-apply(allCauses2, 1, causes)

allCauses4<-sapply(allCauses3, function(x) subset(x, nchar(as.character(x)) <=3))
allCauses4<-sapply(allCauses4, function(x) sub( '(?<=.{1})','0', x, perl=TRUE))

allCauses3<-Map(c, allCauses3, allCauses4)


gbdCauses<-list(allCauses3[c(1)], allCauses3[c(2:7)])
gbdCauses2=setNames(gbdCauses, substr(list.files(pattern="*.csv"),1,nchar(causefiles)-4))

gbdCausesicd10<-lapply(gbdCauses2, rapply, f=c)

#Execute if we want disease icd codes in dataframe
#gbdCVD<-matrix(unlist(gbdCauses3[["cvd"]]), byrow=TRUE, ncol=length(gbdCauses3[1]))

rm(allCauses, allCauses2, allCauses3, allCauses4, allCauses5, gbdCauses, gbdCauses2,newCodes, causes, transposer)
save(gbdCausesicd10, file="R:/working/preparing data/DeathData/cvdCausesicd10.RData")

```



