---
title: "PrepareCVDCauses_ICD9"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Prepare ICD codes for 21 GBD causes                                              
             
## Load libraries, import data, and define attributes 

```{r}
library(stringr)
library(abind)
library(data.table)
library(dplyr)
library(stringi)
dir<-"R:/working/preparing data/DeathData/subsetcvd/icd9/"
setwd(dir)
```

## Prepare ICD codes for 21 GBD causes

```{r}
causefiles=(list.files(pattern="*.csv"))
causefiles=setNames(causefiles, substr(list.files(pattern="*.csv"),1,nchar(causefiles)-4))
allCauses<-lapply(causefiles,read.table,header=F,sep=",", stringsAsFactors=F)
transposer<-function(df){
  z<-t(df)
}
allCauses<-lapply(allCauses, transposer)

#Applied to just ICD9 codes with just numbers
causes2=function(x,output){
  V1 = x[1]
  V2 = word(V1,1,sep="\\-")
  #gsub(".", ".0", V2, invert=)
  V3 = sub("^.+-","", V1)
  seq<-as.data.frame((t(seq(V2,V3, by=0.1))))
  seq2<-seq[which(!stri_sub(as.character(seq),-2,-2)==".")]
  seq2<-paste0(seq2, "0")
  seq2<-seq2[which(!stri_sub(as.character(seq2),1,1)=="0")]
  if ("0" %in% substr(as.character(V2), 1,1)) {
  seq2 = sub("^","0", seq2)}
  seq2<-matrix(unlist(seq2))
}

causes3=function(x,output){
  V1 = x[1]
  V2 = word(V1,1,sep="\\-")
  V3 = sub("^.+-","", V1)
  seq<-as.data.frame((t(seq(V2,V3, by=0.1))))
  if ("0" %in% substr(as.character(V2), 1,1)) {
  seq = sub("^","0", seq)}
  final = gsub("\\.","",seq)
}

allCauses2<-do.call(rbind.data.frame, allCauses)
allCauses3<-apply(allCauses2, 1, causes2)
allCauses4<-apply(allCauses2, 1, causes3)
allCauses5<-Map(c, allCauses3, allCauses4)



  
#Applied to just ICD codes with both letters and numbers
causes=function(x,output){
  V1 = x[1]
  V2 = gsub("[A-Z]","", V1)
  V3 = substring(V1,1,1)
  V4 = word(V2,1,sep="\\-")
  V5 = sub("^.+-","", V2)
  seq<-as.data.frame((t(seq(V4,V5, by=0.1))))
  final = sub("^",V3, seq)
  final = gsub("\\.","",final)
}


gbdicd9<-list(allCauses5)
gbdicd9<-unlist(gbdicd9, recursive=FALSE)

gbdicd9.2<-list(gbdicd9[c(1)], gbdicd9[c(2:4)]) 

causefiles=(list.files(pattern="*.csv"))
gbdCauses2=setNames(gbdicd9.2, substr(causefiles,1,nchar(causefiles)-4))

gbdCausesicd9<-lapply(gbdCauses2, rapply, f=c)

#Execute if we want disease icd codes in dataframe
#gbdAll9<-matrix(unlist(gbdCausesicd9))

rm(allCauses, allCauses2, allCauses3, allCauses4, allCauses5, allCauses6, gbdCauses2, gbdicd9, gbdicd9.2, newCodes, newCodes2, causes, causes2, newnames, transposer)

save(gbdCausesicd9, file="R:/working/preparing data/DeathData/icd9/cvdCausesicd9.RData")
```

