---
title: "map_LE_ggplot"
output: html_document
---


   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, data, and define theme of map                                                     
             (B) Define year, sex, and define map features                                                                
             (C) Map the data and define colours, titles, and legends                                                                        


#Load libraries, data, and define theme of map

```{r}

library(ggplot2)
library(viridis)
library(rgdal)
library(raster)
library(dplyr)

dir<-"R:/working/modeling/temp_dir/plot"
dir2<-"R:/working/preparing data/"
dir3<-"R:/working/modeling/shape_files/shape-mapping"
dir4<-"R:/working/results/"
file<-".rdata"
CTs<-368
year<-c(2016)

#Define the theme for the maps

theme_map<-function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "Ubuntu Regular", color = "#22211d"),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
  panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
  panel.grid.minor = element_blank(),
  plot.background = element_rect(fill = "#f5f5f2", color=NA),
  panel.background = element_rect(fill = "#f5f5f2", color = NA),
  legend.background = element_rect(fill = "#f5f5f2", color = NA),
  panel.border= element_blank()
  )
}

```

#Define year, sex, and define map features

```{r}
#load LE data
#temp for one:
year<-2016
Sex<-1
title<-"Male"

  load(paste0(dir,"/lt_est_", year, file)) #LOAD LE Estimates
  load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
  load(paste0(dir3, CTs, "CTs", file)) #LOAD CT MAP
  shape<-spTransform(shape, crs("+init=epsg:26910"))
  
  PopCTs<-CT %>%
    rename(CTNAME=CT)

  #Prep LE
  birth_all<-est %>%
    filter(sex==Sex & age==0) %>%
    select(area,ex_mean) %>%
    rename(CTNAME = area) %>%
    left_join(PopCTs, by="CTNAME") %>%
    select(CTNAME,ex_mean)
  
  
  #merge shapefiles with LE data
  shapeEXP<-merge(x=shape, y=birth_all, by="CTNAME")
  shapeEXP@data$id = rownames(shape@data)
  shapeEXP$id<-as.numeric(as.character(shapeEXP$id))
  
  
  #Prep and fortify shapefile
  shape@data$id = rownames(shape@data)
  shape_fortified<-fortify(shape, region = "id") %>%
    mutate(id = as.numeric(id))
  shape_data<- shape_fortified %>% left_join(shapeEXP@data, by="id")
  
  #define breaks
  
  breaks<-c(75,78,81,84,87,90)
  min<-min(birth_all$ex_mean)
  max<-max(birth_all$ex_mean)
  

  labels<-c()
  brks<-c(min, breaks, max)
  
  for (i in 1:length(brks)) {
    labels <-c(labels,round(brks[i + 1], 2))
  }
  labels<-labels[1:length(labels)-1]
  
  shape_data$brks<-cut(shape_data$ex_mean, 
                    breaks=brks,
                    labels=labels,
                    include.lowest = TRUE)

  brks_scale<-levels(shape_data$brks)
  labels_scale<-c(paste0(""), 90, 87, 84, 81, 78, 75)

```

#Map the data and define colours, titles, and legends

```{r}

p<-ggplot() +
  geom_polygon(data=shape_data, aes(fill=brks, 
                                  x=long, 
                                  y=lat, 
                                  group=group)) + 
  #municipality outline
  geom_path(data = shape_data, aes (x = long,
                               y = lat,
                               group = group),
            color = "#bababa", size = 0.001) +
  coord_equal() +
  theme_map() +
  theme(
    legend.position = "top",
    legend.text.align = 0,
    legend.background = element_rect(fill = alpha('white', 0.0)),
    legend.text = element_text(size = 10, hjust = 0, color = "#4e4d47"),
    plot.title = element_text(hjust = 0.5, color = "#4e4d47"),
    plot.subtitle = element_text(hjust = 0.5, color = "#4e4d47",
                                 margin = margin(b= -0.1,
                                                 t = -0.1,
                                                 l = 2,
                                                 unit = "cm"),
                                 debug = F),
    legend.title = element_text(size = 10),
    plot.margin = unit(c(.5,.5,.2,.5), "cm"),
    panel.spacing = unit(c(-.1, 0.2,.2,0.2), "cm"),
    panel.border=element_blank(),
    plot.caption=element_text(size =9,
                              hjust = 1,
                              margin = margin(t=0,
                                              b=0.2,
                                              unit = "cm"),
                              color = "#939184")
  ) +
  labs(x=NULL,
       y=NULL,
       title = paste0(year, " Census Tract ", title," Life Expectancy in Metro Vancouver"),
       subtitle = "",
       caption = "Data Source: Vital Statistics, Statistics Canada, PopData BC") 
  #colourbar for discrete data
final<-p +
   scale_fill_manual(
    values=rev(viridis(7)),
    breaks = rev(brks_scale),
    name = "Life expectancy at birth",
    drop=F,
    labels = labels_scale,
    guide=guide_legend(
      direction="horizontal",
      keyheight = unit(3, units = "mm"),
      keywidth = unit(80/length(labels), units = "mm"),
      title.position="top",
      title.hjust = 0.5,
      label.hjust=1,
      nrow=1,
      byrow=T,
      reverse = T,
      label.position = "bottom"
    )
  )



```



