---
title: "map_cause_spplot"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load files and define cause, year, CTs, and directories for mapping                                                     
             (B) Map for male-only, cause-specific mortality data                                                             
             (C) Map for female-only, cause-specific mortality data
             (D) Loop to map both male and female-only, multiple cause-specific mortality data

#Load files and define cause, year, CTs, and directories for mapping

```{r}
library(tidyr)
library(sp)
library(raster)
library(dplyr)
library(viridis)

load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)

causenames<-"prostate"
dir<-"R:/working/modeling/temp_dir/causes/"
file<-".rdata"
CTs<-368
dir2<-"R:/working/preparing data/"
dir3<-"R:/working/modeling/shape_files/shape-mapping"
years<-c(1991)
dir4<-"R:/working/results/"
```

#Map for male-only, cause-specific mortality data

```{r}
#Males

for (this_cause in causenames) { #do not run if needed to only to run one cause - start
     
  for (year in years) {    
  load(paste0(dir,this_cause, "/lt_est_", year, "_males_", CTs, file))
  
  totalsBoth<-totalsMales
    
  load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
  load(paste0(dir3, CTs, "CTs", file)) #LOAD CT SHAPEFILE
  shape<-spTransform(shape, crs("+init=epsg:4326"))###do not run if only running for one cause - finish
  
  #################BOTH########################
    
    range<-round(quantile(totalsBoth$wt_mx_mean, c(.1,.3,.5,.7,.9), na.rm=TRUE), digits=1) ##change sex if needed
  
    #Prepare data to be mapped and plotted
    #at<-c(range)
    at=c(round(quantile(totalsBoth$wt_mx_mean, .05, na.rm=T)),range,round(quantile(totalsBoth$wt_mx_mean, .95, na.rm=T)))
    fifthperc<-round(quantile(totalsBoth$wt_mx_mean, .05, na.rm=T), digits=1)
    ninetyfifthperc<-round(quantile(totalsBoth$wt_mx_mean, .95, na.rm=T), digits=1)
    
    labels=c(paste0(""),fifthperc, range, ninetyfifthperc)
    
    colnames(shape@data)[1]<-"CTNAME"
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    
    #Prep LE
    names(CT)<-c("CTNAME","CT")
    birth_all2<-totalsBoth %>%
      merge(CT, by="CT") %>%
      select(CTNAME,wt_mx_mean) %>%
      rename(mx=wt_mx_mean) 
    birth_all2$CTNAME<-as.numeric(birth_all2$CTNAME)

    #merge shapefiles with current Exp and Mortality
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    shapeMX<-merge(x=shape, y=birth_all2, by="CTNAME")
    
    
    pdf(paste0(dir4, this_cause,"/",this_cause ,"_mxmap_",year,"_males.pdf"))
    
    #mapping life expectancy
    names(shapeMX@data)[2]<-"exp"
    shapeMX$exp<-as.numeric(shapeMX$exp)
    shapeMX@data$exp<-cut(shapeMX@data$exp, breaks=c(0,at,Inf))
    print(
    spplot(shapeMX, "exp",  sp.layout=list("sp.polygons",shapeMX),col.regions=rev(viridis(8)), col="transparent", par.settings = list(axis.line=list(col="transparent")), main=paste0(year," Prostate Cancer Mortality Rate per 100,000 for Males"), colorkey=list(labels=list(at=seq(0.5,length(at)+1), labels=labels))))
    
    dev.off() 
  }
}

      
```

#Map for female-only, cause-specific mortality data

```{r}

for (this_cause in causenames) { #do not run if only needing to run for one cause - start
     
  for (year in years) {    
  load(paste0(dir,this_cause, "/lt_est_", year, "_females_", CTs, file))
  
  totalsBoth<-totalsFemales

  load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
  load(paste0(dir3, CTs, "CTs", file)) #LOAD CT SHAPEFILE
  shape<-spTransform(shape, crs("+init=epsg:4326"))###do not run if only needing to run for one cause - finish
  #################BOTH########################
    
    range<-round(quantile(totalsBoth$wt_mx_mean, c(.1,.3,.5,.7,.9), na.rm=TRUE), digits=1) #change sex if needed
  
    #Prepare data to be mapped and plotted
    at=c(round(quantile(totalsBoth$wt_mx_mean, .05, na.rm=T)),range,round(quantile(totalsBoth$wt_mx_mean, .95, na.rm=T)))
    fifthperc<-round(quantile(totalsBoth$wt_mx_mean, .05, na.rm=T), digits=1)
    ninetyfifthperc<-round(quantile(totalsBoth$wt_mx_mean, .95, na.rm=T), digits=1)
    
    labels=c(paste0(""),fifthperc, range, ninetyfifthperc)
    
    colnames(shape@data)[1]<-"CTNAME"
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    
    #Prep LE
    names(CT)<-c("CTNAME","CT")
    birth_all2<-totalsBoth %>%
      merge(CT, by="CT") %>%
      select(CTNAME,wt_mx_mean) %>%
      rename(mx=wt_mx_mean) 
    birth_all2$CTNAME<-as.numeric(birth_all2$CTNAME)

    #merge shapefiles with current Exp and Mortality
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    shapeMX<-merge(x=shape, y=birth_all2, by="CTNAME")
    
    
    pdf(paste0(dir4, this_cause,"/",this_cause ,"_mxmap_",year,"_females.pdf"))
    
    #mapping life expectancy
    names(shapeMX@data)[2]<-"exp"
    shapeMX$exp<-as.numeric(shapeMX$exp)
    shapeMX@data$exp<-cut(shapeMX@data$exp, breaks=c(0,at,Inf))
    print(
    spplot(shapeMX, "exp",  sp.layout=list("sp.polygons",shapeMX),col.regions=rev(viridis(8)), col="transparent", par.settings = list(axis.line=list(col="transparent")), main=paste0(year," Prostate Cancer Mortality Rate per 100,000 for Females"), colorkey=list(labels=list(at=seq(0.5,length(at)+1), labels=labels))))
    
    dev.off() 
  }
}

```

#Loop to map both male and female-only, multiple cause-specific mortality data

```{r}
for (this_cause in causenames) {

     
  for (year in years) {
        
  load(paste0(dir,this_cause, "/lt_est_", year, "_", CTs, file))
  load(paste0(dir,this_cause, "/lt_est_", year, "_females_", CTs, file))
  load(paste0(dir,this_cause, "/lt_est_", year, "_males_", CTs, file))
  
  load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
  load(paste0(dir3, CTs, "CTs", file)) #LOAD CT SHAPEFILE
  shape<-spTransform(shape, crs("+init=epsg:4326"))
  
  ##FEMALES
    
    range<-round(quantile(totalsFemales$wt_mx_mean, c(.1,.3,.5,.7,.9), na.rm=TRUE), digits=1) ##CHANGE SEX
  
    #Prepare data to be mapped and plotted
    #at<-c(range)
    at=c(round(quantile(totalsFemales$wt_mx_mean, .05)),range,round(quantile(totalsFemales$wt_mx_mean, .95)))
    fifthperc<-round(quantile(totalsFemales$wt_mx_mean, .05), digits=1)
    ninetyfifthperc<-round(quantile(totalsFemales$wt_mx_mean, .95), digits=1)
    
    labels=c(paste0(""),fifthperc, range, ninetyfifthperc)
    
    colnames(shape@data)[1]<-"CTNAME"
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    
    #Prep LE
    names(CT)<-c("CTNAME","CT")
    birth_all2<-totalsFemales %>%
      merge(CT, by="CT") %>%
      select(CTNAME,wt_mx_mean) %>%
      rename(mx=wt_mx_mean) 
    birth_all2$CTNAME<-as.numeric(birth_all2$CTNAME)

    #merge shapefiles with current Exp and Mortality
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    shapeMX<-merge(x=shape, y=birth_all2, by="CTNAME")
    
    pdf(paste0(dir4, this_cause,"/",this_cause ,"_mxmap_",year,"_females.pdf"))
    
    #mapping life expectancy
    names(shapeMX@data)[2]<-"exp"
    shapeMX$exp<-as.numeric(shapeMX$exp)
    shapeMX@data$exp<-cut(shapeMX@data$exp, breaks=c(0,at,Inf))
    print(
    spplot(shapeMX, "exp",  sp.layout=list("sp.polygons",shapeMX), col.regions=rev(viridis(8)), col="transparent", par.settings = list(axis.line=list(col="transparent")), main=paste0(year," ", this_cause, " Mortality Rate per 100,000 for Females"), colorkey=list(labels=list(at=seq(0.5,length(at)+1), labels=labels))))
    
    dev.off()
    

  load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
  load(paste0(dir3, CTs, "CTs", file)) #REPLACE DESIRED CT MAP
  shape<-spTransform(shape, crs("+init=epsg:3857"))
  
  ##MALES
    
    range<-round(quantile(totalsMales$wt_mx_mean, c(.1,.3,.5,.7,.9), na.rm=TRUE), digits=1) ##CHANGE SEX
  
    #Prepare data to be mapped and plotted
    at=c(round(quantile(totalsMales$wt_mx_mean, .05)),range,round(quantile(totalsMales$wt_mx_mean, .95)))
    fifthperc<-round(quantile(totalsMales$wt_mx_mean, .05), digits=1)
    ninetyfifthperc<-round(quantile(totalsMales$wt_mx_mean, .95), digits=1)
    
    labels=c(paste0(""),fifthperc, range, ninetyfifthperc)
    
    colnames(shape@data)[1]<-"CTNAME"
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    
    #Prep LE
    names(CT)<-c("CTNAME","CT")
    birth_all2<-totalsMales %>%
      merge(CT, by="CT") %>%
      select(CTNAME,wt_mx_mean) %>%
      rename(mx=wt_mx_mean) 
    birth_all2$CTNAME<-as.numeric(birth_all2$CTNAME)

    #merge shapefiles with current Exp and Mortality
    shape$CTNAME<-as.numeric(as.character(shape$CTNAME))
    shapeMX<-merge(x=shape, y=birth_all2, by="CTNAME")
    
    
    pdf(paste0(dir4, this_cause,"/",this_cause ,"_mxmap_",year,"_males.pdf"))
    
    #mapping life expectancy
    names(shapeMX@data)[2]<-"exp"
    shapeMX$exp<-as.numeric(shapeMX$exp)
    shapeMX@data$exp<-cut(shapeMX@data$exp, breaks=c(0,at,Inf))
    print(
    spplot(shapeMX, "exp",  sp.layout=list("sp.polygons",shapeMX), col.regions=rev(viridis(8)), col="#bababa", par.settings = list(axis.line=list(col="transparent")), main=paste0(year," ", this_cause, " Mortality Rate per 100,000 for Males"), colorkey=list(labels=list(at=seq(0.5,length(at)+1), labels=labels))))
    
    dev.off()
 
    load(paste0(dir2, "CT", CTs, file))#LOAD CT FILE
    load(paste0(dir3, CTs, "CTs", file)) #REPLACE DESIRED CT MAP
    shape<-spTransform(shape, crs("+init=epsg:4326"))
  }
}
```

