---
title: "adjustPop"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Create dummy pop file that includes all rows of data                                                   
             (B) Load crosswalked population file, summarise data, and relabel columns as needed                                                      
             (C) Separately interpolate CT 58 -only needed to run once                                                                        
             (D) Linearly interpolate CTs 161.07, 251.01, 290.06, 400.07, 400.08
             (E) Linearly interpolate
             (F) Remove interpolated CTs removed from original dataset
             (G) Finalise columns, summarise, and save interpolate pop file
             (H)
             (I) Load merge file and merge popfiles as needed for shapefiles
  
             
#Create dummy pop file that includes all rows of data 

```{r}
age<-c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85)
sex<-c(1,2)
pop<-c(0)
year<-c(1991,1996,2001,2006,2011,2016)

popDF<-data.frame(age,pop)

popDF2 <- cbind(rep(1, nrow(popDF)), popDF)
for(i in 2){
  a <- cbind(rep(i, nrow(popDF)), popDF)
  names(a) <- names(popDF2)
  popDF2 <- rbind(popDF2, a)
}
popDF2 <- as.data.frame(popDF2)
names(popDF2)[1]<-"sex"

years<-c(1996,2001,2006,2011,2016)

b <- cbind(rep(1991, nrow(popDF2)), popDF2)
for(i in years){
  a <- cbind(rep(i, nrow(popDF2)), popDF2)
  names(a) <- names(b)
  b <- rbind(b, a)
}
popDF <- as.data.frame(b)
names(popDF)[1]<-"year"


load("R:/working/preparing data/listCTs.RData")
PopUniqueCT<-as.data.frame(listCTs)
PopUniqueCT<-PopUniqueCT[-1,]

b <- cbind(rep(1.01, nrow(popDF)), popDF)
for(i in PopUniqueCT){
  a <- cbind(rep(i, nrow(popDF)), popDF)
  names(a) <- names(b)
  b <- rbind(b, a)
}
popDF <- as.data.frame(b)
names(popDF)[1]<-"CT"

popDF<-popDF %>%
  select(year,CT,sex,age,pop)

setDT(popDF)

#remove all but my popDF file
rm(list=(ls()[ls()!="popDF"]))

```

#Load crosswalked population file, summarise data, and relabel columns as needed

```{r}

library(dplyr)

load("R:/working/preparing data/PopulationData/prepareCrosswalk/censusPop.RData")

pop<-pop %>% 
  drop_na(CT) %>%
  group_by(year,CT,sex,age) %>%
  summarise(pop=mean(pop))

popDF2 <- popDF %>%
  merge(pop, by=c("year","CT","sex","age"))

popDF2$pop.x<-popDF2$pop.y

pop<-popDF2 %>%
  select(year,CT,sex,age,pop.x) %>%
  rename(pop=pop.x)

#remove all but my popDF file
rm(list=(ls()[ls()!="pop"]))

pop91<-subset(pop, year==1991)

pop96<-subset(pop, year==1996)

pop01<-subset(pop, year==2001)

pop06<-subset(pop, year==2006)

pop11<-subset(pop, year==2011)

pop16<-subset(pop, year==2016)


pop91<-select(pop91,2,3,4,5)
pop96<-select(pop96,2,3,4,5)
pop01<-select(pop01,2,3,4,5)
pop06<-select(pop06,2,3,4,5)
pop11<-select(pop11,2,3,4,5)
pop16<-select(pop16,2,3,4,5)

names(pop91)[4]<-"pop91"
names(pop96)[4]<-"pop96"
names(pop01)[4]<-"pop01"
names(pop06)[4]<-"pop06"
names(pop11)[4]<-"pop11"
names(pop16)[4]<-"pop16"

```


#Separately interpolate CT 58 -only needed to run once

```{r}
pop06.58<-subset(pop06, CT==58.00)
pop16.58<-subset(pop16, CT==58.00)

#1990-1996
pop58<-cbind(pop06.58, pop16.58)
pop58<-pop58[,c(1,2,3,4,8)]
diff58<-(pop58$pop16-pop58$pop06)/10

pop58$pop07<-pop58$pop06+diff58 
pop58$pop08<-pop58$pop07+diff58
pop58$pop09<-pop58$pop08+diff58
pop58$pop10<-pop58$pop09+diff58
pop58$pop11<-pop58$pop10+diff58
pop58$pop12<-pop58$pop11+diff58
pop58$pop13<-pop58$pop12+diff58
pop58$pop14<-pop58$pop13+diff58
pop58$pop15<-pop58$pop14+diff58


pop58.7<-pop58[,c("CT","sex","age","pop07")]
pop58.8<-pop58[,c("CT","sex","age","pop08")]
pop58.9<-pop58[,c("CT","sex","age","pop09")]
pop58.10<-pop58[,c("CT","sex","age","pop10")]
pop58.11<-pop58[,c("CT","sex","age","pop11")]
pop58.12<-pop58[,c("CT","sex","age","pop12")]
pop58.13<-pop58[,c("CT","sex","age","pop13")]
pop58.14<-pop58[,c("CT","sex","age","pop14")]
pop58.15<-pop58[,c("CT","sex","age","pop15")]

names(pop06.58)<-c("CT","sex","age","pop")
names(pop58.7)<-c("CT","sex","age","pop")
names(pop58.8)<-c("CT","sex","age","pop")
names(pop58.9)<-c("CT","sex","age","pop")
names(pop58.10)<-c("CT","sex","age","pop")
names(pop58.11)<-c("CT","sex","age","pop")
names(pop58.12)<-c("CT","sex","age","pop")
names(pop58.13)<-c("CT","sex","age","pop")
names(pop58.14)<-c("CT","sex","age","pop")
names(pop58.15)<-c("CT","sex","age","pop")
names(pop16.58)<-c("CT","sex","age","pop")

pop06.58$year<-2006
pop58.7$year<-2007
pop58.8$year<-2008
pop58.9$year<-2009
pop58.10$year<-2010
pop58.11$year<-2011
pop58.12$year<-2012
pop58.13$year<-2013
pop58.14$year<-2014
pop58.15$year<-2015
pop16.58$year<-2016


pop58.final<-rbind(pop06.58,pop58.7,pop58.8,pop58.9,pop58.10,pop58.11,pop58.12,pop58.13,pop58.14,pop58.15,pop16.58)


```

#Linearly interpolate CTs 161.07, 251.01, 290.06, 400.07, 400.08 - not necessary to re-run unless you want to modify how to deal with certain CTs that don't interpolate well into other years - 290.06, 400.07, and 400.08

```{r}
pop11.161<-subset(pop06, CT==161.07)
pop16.161<-subset(pop11, CT==161.07)
pop161<-cbind(pop11.161, pop16.161)


pop11.251<-subset(pop06, CT==251.01)
pop16.251<-subset(pop11, CT==251.01)
pop251<-cbind(pop11.251, pop16.251)


pop11.290<-subset(pop06, CT==290.07)
pop16.290<-subset(pop11, CT==290.07)
pop290<-cbind(pop11.290, pop16.290)


pop11.4007<-subset(pop06, CT==400.07)
pop16.4007<-subset(pop11, CT==400.07)
pop4007<-cbind(pop11.4007, pop16.4007)


pop11.4008<-subset(pop06, CT==400.08)
pop16.4008<-subset(pop11, CT==400.08)
pop4008<-cbind(pop11.4008, pop16.4008)


listPop<-list(pop161,pop251,pop290,pop4007,pop4008)

listpop<-lapply(listPop, function(x) x[,c(1,2,3,4,8)])
for(i in seq_along(listpop)){
  listpop[[i]]$diff <-listpop[[i]]$pop11-listpop[[i]]$pop06/5
}
for(i in seq_along(listpop)){
  listpop[[i]]$pop12 <-listpop[[i]]$pop11+listpop[[i]]$diff
  listpop[[i]]$pop13 <-listpop[[i]]$pop12+listpop[[i]]$diff
  listpop[[i]]$pop14 <-listpop[[i]]$pop13+listpop[[i]]$diff
  listpop[[i]]$pop15 <-listpop[[i]]$pop14+listpop[[i]]$diff
  listpop[[i]]$pop16 <-listpop[[i]]$pop15+listpop[[i]]$diff
}

listpop<-lapply(listpop, function(x) x[,!c(4:6)])

popMissing<-do.call(rbind, lapply(listpop, function(x) x[,1:8]))

popMissing12<-popMissing[,c("CT","sex","age","pop12")]
popMissing13<-popMissing[,c("CT","sex","age","pop13")]
popMissing14<-popMissing[,c("CT","sex","age","pop14")]
popMissing15<-popMissing[,c("CT","sex","age","pop15")]
popMissing16<-popMissing[,c("CT","sex","age","pop16")]

names(popMissing12)[4]<-"pop"
names(popMissing13)[4]<-"pop"
names(popMissing14)[4]<-"pop"
names(popMissing15)[4]<-"pop"
names(popMissing16)[4]<-"pop"

popMissing12$year<-2012
popMissing13$year<-2013
popMissing14$year<-2014
popMissing15$year<-2015
popMissing16$year<-2016

popMissing<-rbind(popMissing12,popMissing13,popMissing14,popMissing15,popMissing16)
popMissing<-popMissing[,c("year","CT","sex","age","pop")]

popMissing$pop[which(popMissing$pop<0)]<-0 #we imputed 0 in ages of areas where there were 0, another option is to impute the 2011 estimates

popMissings<-rbind(popMissing,pop58.final)

save(popMissings, file="R:/working/preparing data/PopulationData/popMissings.RDATA")

```


#linearly interpolate

```{r}
#1990-1996
pop9691<-cbind(pop96, pop91)
pop9691<-pop9691[,c(1,2,3,4,8)]
diff9691<-(pop9691$pop96-pop9691$pop91)/5

pop9691$pop90<-pop9691$pop91 ##left at population at 1991
pop9691$pop92<-pop9691$pop91+diff9691
pop9691$pop93<-pop9691$pop92+diff9691
pop9691$pop94<-pop9691$pop93+diff9691
pop9691$pop95<-pop9691$pop94+diff9691

pop90<-pop9691[,c("CT","sex","age","pop90")]
pop92<-pop9691[,c("CT","sex","age","pop92")]
pop93<-pop9691[,c("CT","sex","age","pop93")]
pop94<-pop9691[,c("CT","sex","age","pop94")]
pop95<-pop9691[,c("CT","sex","age","pop95")]

#1996-2001
pop0196<-cbind(pop01,pop96)
pop0196<-pop0196[,c(1,2,3,4,8)]
diff0196<-(pop0196$pop01-pop0196$pop96)/5

pop0196$pop97<-pop0196$pop96+diff0196
pop0196$pop98<-pop0196$pop97+diff0196
pop0196$pop99<-pop0196$pop98+diff0196
pop0196$pop00<-pop0196$pop99+diff0196

pop97<-pop0196[,c("CT","sex","age","pop97")]
pop98<-pop0196[,c("CT","sex","age","pop98")]
pop99<-pop0196[,c("CT","sex","age","pop99")]
pop00<-pop0196[,c("CT","sex","age","pop00")]

#2001-2006
pop0106<-cbind(pop01,pop06)
pop0106<-pop0106[,c(1,2,3,4,8)]
diff0601<-(pop0106$pop06-pop0106$pop01)/5

pop0106$pop02<-pop0106$pop01+diff0601
pop0106$pop03<-pop0106$pop02+diff0601
pop0106$pop04<-pop0106$pop03+diff0601
pop0106$pop05<-pop0106$pop04+diff0601

pop02<-pop0106[,c("CT","sex","age","pop02")]
pop03<-pop0106[,c("CT","sex","age","pop03")]
pop04<-pop0106[,c("CT","sex","age","pop04")]
pop05<-pop0106[,c("CT","sex","age","pop05")]

#2006-2011
pop0611<-cbind(pop06,pop11)
pop0611<-pop0611[,c(1,2,3,4,8)]
diff1106<-(pop0611$pop11-pop0611$pop06)/5

pop0611$pop07<-pop0611$pop06+diff1106
pop0611$pop08<-pop0611$pop07+diff1106
pop0611$pop09<-pop0611$pop08+diff1106
pop0611$pop10<-pop0611$pop09+diff1106

pop07<-pop0611[,c("CT","sex","age","pop07")]
pop08<-pop0611[,c("CT","sex","age","pop08")]
pop09<-pop0611[,c("CT","sex","age","pop09")]
pop10<-pop0611[,c("CT","sex","age","pop10")]

#2011-2016
pop1116<-cbind(pop16,pop11)
pop1116<-pop1116[,c(1,2,3,4,8)]
diff1611<-(pop1116$pop16-pop1116$pop11)/5

pop1116$pop12<-pop1116$pop11+diff1611
pop1116$pop13<-pop1116$pop12+diff1611
pop1116$pop14<-pop1116$pop13+diff1611
pop1116$pop15<-pop1116$pop14+diff1611

pop12<-pop1116[,c("CT","sex","age","pop12")]
pop13<-pop1116[,c("CT","sex","age","pop13")]
pop14<-pop1116[,c("CT","sex","age","pop14")]
pop15<-pop1116[,c("CT","sex","age","pop15")]

rm(pop,pop0106,pop1116,pop0196,pop9691,pop0611,diff0196,diff0601,diff1106,diff1611,diff9691)

```

#Remove interpolated CTs removed from original dataset

```{r}
pop06<-subset(pop06, CT!=58)
pop07<-subset(pop07, CT!=58)
pop08<-subset(pop08, CT!=58)
pop09<-subset(pop09, CT!=58)
pop10<-subset(pop10, CT!=58)
pop11<-subset(pop11, CT!=58 & CT!=161.07 & CT!=251.01 &  CT!=290.06 &  CT!=400.07 &  CT!=400.08)
pop12<-subset(pop12, CT!=58 & CT!=161.07 & CT!=251.01 &  CT!=290.06 &  CT!=400.07 &  CT!=400.08)
pop13<-subset(pop13, CT!=58 & CT!=161.07 & CT!=251.01 &  CT!=290.06 &  CT!=400.07 &  CT!=400.08)
pop14<-subset(pop14, CT!=58 & CT!=161.07 & CT!=251.01 &  CT!=290.06 &  CT!=400.07 &  CT!=400.08)
pop15<-subset(pop15, CT!=58 & CT!=161.07 & CT!=251.01 &  CT!=290.06 &  CT!=400.07 &  CT!=400.08)
pop16<-subset(pop16, CT!=58 & CT!=161.07 & CT!=251.01 & CT!=290.06 &  CT!=400.07 &  CT!=400.08)
```

#Finalise columns, summarise, and save interpolate pop file

```{r}


pop<-lapply(ls(), get)
colnames<-c("CT","sex","age", "pop")
years<-list(2000,2001,2002,2003,2004,2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,1990,1991,1992,1993,1994,1995,1996,1997,1998,1999)

pop<-lapply(pop, setNames, colnames)
pop<-Map(cbind, pop, year=years)

popTest<-do.call(rbind, lapply(pop, function(x) x[,1:5]))
pop<-popTest[,c(5,1,2,3,4)]

load("R:/working/preparing data/PopulationData/popMissings.RDATA")

pop_final<-rbind(pop,popMissings)
pop_final[is.na(pop_final)]<-0
pop<-subset(pop_final,CT!=270 & CT>0)

pop <-pop %>%
    group_by(year,CT,sex,age) %>%
    summarise(pop=mean(pop))

popInfo <-pop %>%
    group_by(year,CT,) %>%
    tally() #double checking 36 age groups present in all years and CT combos

pop<-pop %>%
  arrange(CT)
pop<-setDT(pop)

save(pop, file="R:/working/preparing data/PopulationData/pop_interpolated.RDATA")
```


##Load merge file and merge popfiles as needed for shapefiles

```{r}
merge<-read.csv("R:/working/preparing data/Merge/merge_all12.csv", header=T)

for (i in 1:nrow(merge)){
  pop$CT[which(pop$CT==as.numeric(merge[i,1]))]<-as.numeric(merge[i,2])
}

pop <-pop %>%
  group_by(CT,year,sex,age) %>%
  summarise(pop=sum(pop)) %>%
  mutate(pop=round(pop))

pop<-pop %>%
  group_by(CT)
pop$CT<-as.numeric(as.factor(pop$CT))
pop$CT<-(pop$CT-1)
pop$CT<-as.integer(pop$CT)
pop<-setDT(pop)
length(unique(pop$CT))

#Finalise and save pop file

save(pop, file="R:/working/modeling/pop_files/pop_new_368CTs.RData")

```

