---
title: "PreparePop"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and data                                                     
             (B) Relabel all years to appropriate years                                                                
             (C) Merge all datasets into one                                                                        
             (D) Subset population files for each census year
             (E) Load crosswalk data
             (F) Apply crosswalk tables to population data
             (G) Extract some descriptive statistics
             (H) Relabel some columns and save crosswalked population files

             
## Load the relevant libraries and data files

```{r}
library("dplyr")
library("tidyr")
library("data.table")
dir<-"R:/working/preparing data/PopulationData/allyears/"
year<-1990:2016
census<-c(2016,2011,2006,2001,1996,1991)

for (i in year) {
  load(paste0(dir, i, "/allPop", i, ".Rdata"))
}

```


## Relabel all years to appropriate years

```{r}
Pop1990$year <-rep(1990,nrow(Pop1990))
Pop1991$year <-rep(1991,nrow(Pop1991))
Pop1992$year <-rep(1992,nrow(Pop1992))
Pop1993$year <-rep(1993,nrow(Pop1993))
Pop1994$year <-rep(1994,nrow(Pop1994))
Pop1995$year <-rep(1995,nrow(Pop1995))
Pop1996$year <-rep(1996,nrow(Pop1996))
Pop1997$year <-rep(1997,nrow(Pop1997))
Pop1998$year <-rep(1998,nrow(Pop1998))
Pop1999$year <-rep(1999,nrow(Pop1999))
Pop2000$year <-rep(2000,nrow(Pop2000))
Pop2001$year <-rep(2001,nrow(Pop2001))
Pop2002$year <-rep(2002,nrow(Pop2002))
Pop2003$year <-rep(2003,nrow(Pop2003))
Pop2004$year <-rep(2004,nrow(Pop2004))
Pop2005$year <-rep(2005,nrow(Pop2005))
Pop2006$year <-rep(2006,nrow(Pop2006))
Pop2007$year <-rep(2007,nrow(Pop2007))
Pop2008$year <-rep(2008,nrow(Pop2008))
Pop2009$year <-rep(2009,nrow(Pop2009))
Pop2010$year <-rep(2010,nrow(Pop2010))
Pop2011$year <-rep(2011,nrow(Pop2011))
Pop2012$year <-rep(2012,nrow(Pop2012))
Pop2013$year <-rep(2013,nrow(Pop2013))
Pop2014$year <-rep(2014,nrow(Pop2014))
Pop2015$year <-rep(2015,nrow(Pop2015))
```

## Merge all datasets into one

```{r}
popfiles<-rbind(Pop1990,Pop1991,Pop1992,Pop1993,Pop1994,Pop1995,Pop1996,Pop1997,Pop1998,Pop1999, Pop2000, Pop2001, Pop2002, Pop2003, Pop2004, Pop2005, Pop2006, Pop2007, Pop2008, Pop2009, Pop2010, Pop2011, Pop2012, Pop2013, Pop2014, Pop2015, Pop2016)

popfiles<-subset(popfiles, CT>0)

popInfo <- pop %>%
  group_by(year, CT) %>%
  summarise(pop=sum(pop))

popNA<-popfiles[is.na(popfiles$pop),]
popNaCount<-popNA %>%
  group_by(year,CT) %>%
  summarise()

#popfiles[is.na(popfiles)]<-0 #included 0 for areas with NA values
pop<-popfiles

```

## Recode age and save

```{r}
#Merge all datasets
detach("package:dplyr", unload=TRUE)
library("car")
pop$age<-recode(pop$age,"4=0;9=5;14=10;19=15;24=20;29=25;34=30;39=35;44=40;49=45;54=50;59=55;64=60;69=65;74=70;79=75;84=80;85=85")#must unload dplyr for recode in car package to work

library("dplyr")
pop<-subset(pop,CT>0)
```


## Subset population files for each census year

```{r}
popfiles9093<-subset(pop, year == 1991)

popfiles9498<-subset(pop, year == 1996)

popfiles9903<-subset(pop, year == 2001)

popfiles0408<-subset(pop, year == 2006)

popfiles0913<-subset(pop, year == 2011)

popfiles1416<-subset(pop, year == 2016)

names(popfiles9093)[2]<-"ctuid_s"
names(popfiles9498)[2]<-"ctuid_s"
names(popfiles9903)[2]<-"ctuid_s"
names(popfiles0408)[2]<-"ctuid_s"
names(popfiles0913)[2]<-"ctuid_s"
names(popfiles1416)[2]<-"ctuid_s"


library("dplyr")
save(popfiles9093, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop9093.RData")
save(popfiles9498, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop9498.RData")
save(popfiles9903, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop9903.RData")
save(popfiles0408, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop0408.RData")
save(popfiles0913, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop0913.RData")
save(popfiles1416, file="R:/working/preparing data/PopulationData/prepareCrosswalk/pop1416.RData")

```

## Load crosswalk data and pop data if not already loaded


```{r}

#load crosswalks
crosswalk91<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk91.csv", header=TRUE)
crosswalk96<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk96.csv", header=TRUE)
crosswalk01<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk01.csv", header=TRUE)
crosswalk06<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk06.csv", header=TRUE)
crosswalk11<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk11.csv", header=TRUE)


#load popfiles
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop9093.RData")
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop9498.RData")
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop0408.RData")
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop0913.RData")
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop1416.RData")
load("R:/working/preparing data/PopulationData/prepareCrosswalk/pop9903.RData")
```

## Apply crosswalk tables to population data

```{r}
#Multiply 1991 pop file with crosswalk table to get 2016 CT estimates
pop9093<-merge(x=crosswalk91, y=popfiles9093, by="ctuid_s", all.y=TRUE)
pop9093<-subset(pop9093, ctuid_s>0)
pop9093<-pop9093[order(pop9093$ctuid_s,pop9093$year,pop9093$sex, pop9093$age),]
pop9093<-pop9093 %>% mutate(pop9093 = w * pop)
pop1991 = pop9093 %>% group_by(ctuid_t, year, sex, age) %>%
  summarise(pop9093 = sum(pop9093, na.rm = T))
pop1991<-pop1991[order(pop1991$ctuid_t,pop1991$year,pop1991$sex, pop1991$age),]

#Multiply 1996 pop file with crosswalk table to get 2016 CT estimates
pop9498<-merge(x=crosswalk96, y=popfiles9498, by="ctuid_s", all.y=TRUE)
pop9498<-subset(pop9498, ctuid_s>0)
pop9498<-pop9498[order(pop9498$ctuid_s,pop9498$year,pop9498$sex, pop9498$age),]
pop9498<-pop9498 %>% mutate(pop9498 = w * pop)
pop1996 = pop9498 %>% group_by(ctuid_t, year, sex, age) %>%
  summarise(pop9498 = sum(pop9498, na.rm = T))
pop1996<-pop1996[order(pop1996$ctuid_t,pop1996$year,pop1996$sex, pop1996$age),]

#Multiply 2001 pop file with crosswalk table to get 2016 CT estimates
pop9903<-merge(x=crosswalk01, y=popfiles9903, by="ctuid_s", all.y=TRUE)
pop9903<-subset(pop9903, ctuid_s>0)
pop9903<-pop9903[order(pop9903$ctuid_s,pop9903$year,pop9903$sex, pop9903$age),]
pop9903<-pop9903 %>% mutate(pop9903 = w * pop)
pop2001 = pop9903 %>% group_by(ctuid_t, year, sex, age) %>%
  summarise(pop9903 = sum(pop9903, na.rm = T))
pop2001<-pop2001[order(pop2001$ctuid_t,pop2001$year,pop2001$sex, pop2001$age),]

#Multiply 2006 pop file with crosswalk table to get 2016 CT estimates
pop0408<-merge(x=crosswalk06, y=popfiles0408, by="ctuid_s", all.y=TRUE)
pop0408<-subset(pop0408, ctuid_s>0)
pop0408<-pop0408[order(pop0408$ctuid_s,pop0408$year,pop0408$sex, pop0408$age),]
pop0408<-pop0408 %>% mutate(pop0408 = w * pop)
pop2006 = pop0408 %>% group_by(ctuid_t, year, sex, age) %>%
  summarise(pop0408 = sum(pop0408, na.rm = T))
pop2006<-pop2006[order(pop2006$ctuid_t,pop2006$year,pop2006$sex, pop2006$age),]

#Multiply 2011 pop file with crosswalk table to get 2016 CT estimates
pop0913<-merge(x=crosswalk11, y=popfiles0913, by="ctuid_s", all.y=TRUE)
pop0913<-subset(pop0913, ctuid_s>0)
pop0913<-pop0913[order(pop0913$ctuid_s,pop0913$year,pop0913$sex, pop0913$age),]
pop0913<-pop0913 %>% mutate(pop0913 = w * pop)
pop2011 = pop0913 %>% group_by(ctuid_t, year, sex, age) %>%
  summarise(pop0913 = sum(pop0913, na.rm = T))
pop2011<-pop2011[order(pop2011$ctuid_t,pop2011$year,pop2011$sex, pop2011$age),]

pop2016<-popfiles1416
pop2016<-subset(pop2016, ctuid_s>0)
names(pop2016)[2]<-"ctuid_t"

save(pop1991, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop1991.RData")
save(pop1996, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop1996.RData")
save(pop2001, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop2001.RData")
save(pop2006, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop2006.RData")
save(pop2011, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop2011.RData")
save(pop2016, file="R:/working/preparing data/PopulationData/prepareCrosswalk/newPop2016.RData")

#relabel some columns

pop1991<-data.frame(pop1991)
pop1996<-data.frame(pop1996)
pop2001<-data.frame(pop2001)
pop2006<-data.frame(pop2006)
pop2011<-data.frame(pop2011)
pop2016<-data.frame(pop2016)

names(pop1991)[5]<-"pop"
names(pop1996)[5]<-"pop"
names(pop2001)[5]<-"pop"
names(pop2006)[5]<-"pop"
names(pop2011)[5]<-"pop"
names(pop2016)[5]<-"pop"

pop<-rbind(pop1991,pop1996,pop2001,pop2006,pop2011,pop2016)
names(pop)[1]<-"CT"

```

# Extract some descriptive statistics

```{r}

popInfo <- pop %>%
  group_by(year, CT) %>%
  summarise(pop=sum(pop))

popNA<-pop[is.na(pop$pop),]
popNaCount<-popNA %>%
  group_by(year,CT) %>%
  summarise()

```

#Relabel some columns and save census population file

```{r}
#Order data sequentially 
pop<-pop[order(pop$CT),]

#Change from numeric to integer
pop$sex<-as.integer(pop$sex)

pop$age<-as.numeric(as.character(pop$age))
pop$year<-as.integer(pop$year)

#Convert final data frame into data table
pop<-setDT(pop)

pop<-pop[,c(2,1,3,4,5)]

save(pop, file="R:/working/preparing data/PopulationData/prepareCrosswalk/censusPop.RData")
```

