---
title: "categoriseAllDeaths"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and import data                                                  
             (B) Select years of death
             (C) Recode age variable
             (D) Load and apply crosswalk 
             (E) Relabel, summarise, and save crosswalked death file
             
## Load the relevant libraries

```{r}
library("tidyr")
library("data.table")
library("dplyr")

load("R:/working/preparing data/DeathData/causes/subset/deaths_subset.rdata")
load("R:/working/preparing data/listCTs.RData")
listCTs<-as.data.frame(listCTs)
names(listCTs)<-"ctuid_t"
load("R:/working/preparing data/CTIDs.RData")

```

## Select years of death

```{r}
deaths <- subset(deaths, year>1989)
deaths<-deaths[order(deaths$CT),]

```

## Recode age variable

```{r}
deaths$age<-car::recode(deaths$age,"4=0;9=5;14=10;19=15;24=20;29=25;34=30;39=35;44=40;49=45;54=50;59=55;64=60;69=65;74=70;79=75;84=80;85=85")#must unload dplyr for recode in car package to work
```


## Load and apply crosswalk 

```{r}
crosswalk11<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk11.csv", header=TRUE)
load("R:/working/preparing data/CT2011.RData")
names(CT2011)<-"ctuid_t"

names(deaths)[2]<-"ctuid_s"


#Multiply 2011 death file with crosswalk table to get 2016 CT estimates
death2011<-merge(x=crosswalk11, y=deaths, by="ctuid_s", all.y=TRUE)
death2011<-subset(death2011, ctuid_s>0)
death2011<-death2011[order(death2011$ctuid_s,death2011$year,death2011$sex, death2011$age),]
death2011<-death2011 %>% mutate(death2011 = w * death)
death2016 = death2011 %>% group_by(ctuid_t, year, sex, age) 
death2016<-death2016[order(death2016$ctuid_t,death2016$year,death2016$sex, death2016$age),]

deathUniqueCT2016<-setDT(death2016)[,lapply(.SD, uniqueN), ctuid_t]
#isolate to find missing CTs for each year
death2016m<-select(deathUniqueCT2016, "ctuid_t")
#find missing CTs
missing2016<-anti_join(CT2011, death2016m, by="ctuid_t")

```

## Relabel, summarise, and save crosswalked death file

```{r}
deaths<-data.frame(death2016)
names(deaths)[10]<-"death"

deathNa<-data.frame(table(is.na(deaths)))#663
names(deaths)[3]<-"CT"
deaths<-deaths[,c(6,3,7,8,10)]

save(deaths, file="R:/working/modeling/deaths/deaths_subset.RData")
```
