---
title: "categoriseAllDeaths"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and import data                                                  
             (B) Categorise age groups
             (C) Keep final columns and change variable types
             (D) Count number of deaths and save file
             (E) EXTRA: descriptive info

## Load the relevant libraries

```{r}
library("dplyr")
library("tidyr")
library("data.table")
library(readr)

setwd("R:/working/preparing data/DeathData")
allDeaths<-read.csv(file="final_mv_allyears_sensitivity5_excludeNonVan.csv", header=TRUE, check.names = FALSE)
```

## Categorise age groups

```{r}
#Categorise age groups and rename column names
agebreaks <-c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,113)
agelabels <-c("0","5","10","15","20","25","30","35","40","45","50","55","60","65","70","75","80","85")

deaths<-setDT(allDeaths) [, agegroups := cut(AGE, breaks = agebreaks, right= FALSE, labels = agelabels)]

```

## Keep final columns and change variable types

```{r}

colnames(deaths)[colnames(deaths)=="DEATHDTCCYY"]<-"year"
colnames(deaths)[colnames(deaths)=="agegroups"]<-"age"

deaths$sex<-recode(deaths$SEX, 'F'=2, 'M'=1, 'U'=3)


deaths$sex<-as.numeric(as.character(deaths$sex))
deaths$age<-as.numeric(as.character(deaths$age))

deaths <-dplyr::select(deaths,year,CT,sex,age)

```

## Count number of deaths and save file

```{r}
#Count number of deaths
mortcounts<-count(deaths, year, CT, sex, age)
colnames(mortcounts)[colnames(mortcounts)=="n"]<-"death"

#sort data by sequentially of CT
deaths<-mortcounts[order(mortcounts$CT),]

setDT(deaths)
write.csv(deaths, file="R:/working/preparing data/DeathData/years/allyears/allDeaths.csv")
save(deaths, file="R:/working/preparing data/DeathData/years/final/deaths_restrict.RData")

```


##EXTRA: descriptive info

```{r}
deathTally<-deaths %>%
  group_by(year) %>%
  summarise(deaths=sum(death))

#find death data with missing age information
deathsAgeNA<-deaths[is.na(deaths$age),] #532
deathsAgeNA<-deathsAgeNA %>%
  #group_by(year) %>%
  summarise(deaths=sum(death))
```

