---
title: "prepareRawMortality"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                      
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and import data                                                  
             (B) Create a function that imports mortality dataset and labels the columns
             (C) Loop the import function for all years of data
             (D) Change data from list to data frame
             (E) Keep only PC_CT that are in Vancouver CMA and then remove the first 5 numbers
             (F) Include only CTs and variables of interest and save file

## Load the relevant libraries and import data

```{r}
library("dplyr")
library("tidyr")
library("stringr")
load(file="R:/working/preparing data/CT2011.Rdata")
load(file="R:/working/preparing data/DeathData/df_dataYear.Rdata")
CT2011<-as.matrix(CT2011)
van<-59933
```

## Create a function that imports mortality dataset and labels the columns

```{r}
import <-function(p,filename){
  setwd("R:")
  filepath = paste(p,filename,sep="")
  dat <-read.fwf(file=filepath,fill=TRUE,widths=c(8,4,2,50,3,1,1,1,1,4,2,3,1,50,25,50,3,1,2,5,100,4,2,10,10,12,12,1)) 
  colnames(dat) <-c("REGDT","DEATHDTCCYY","DEATHDTMM","DEATHCITY","DEATHPC_3","SEX","MARITAL","NATIVE","ONRESERVE","BIRTHDTCCYY","BIRTHDTMM","AGE","AGEUNITS","BIRTHCITY","OCCUPATION","CITY","PC_3","OTHER_INFO","ICD_VERSN","CAUSE","AXIS_CODES","DATAYR","VERSION","SEQNO","STUDYID","DEATHPC_6_CT","PC_CT","LINEFEED")

  dat<<-dat
  }

```

## Loop the import function for all years of data

```{r}
p <- "/DATA/2018-09-25/deaths/"
dataFiles<-list.files(path="/DATA/2018-09-25/deaths/",pattern=".B.dat.gz")

dataYear<-list()
for (x in c(1:length(dataFiles))) {
  dataYear[[x]] <- import(p,dataFiles[x])
}
```

## Change data from list to data frame

```{r}
df_dataYear <- as.data.frame(dataYear[[1]])
for (x in c(2:length(dataYear))) {
  df_dataYear <- rbind(df_dataYear, as.data.frame(dataYear[[x]]))
}
```

## Keep only PC_CT that are in Vancouver CMA and then remove the first 5 numbers

```{r}
cities<-c("VANCOUVER", "BURNABY","`WEST VANCOUVER`","`NORTH VANCOUVER`", "BELCARRA", "ANMORE", "`PORT MOODY`", "COQUITLAM", "`PORT COQUITLAM`", "`NEW WESTMINSTER`", "`PITT MEADOWS`", "RICHMOND", "DELTA", "SURREY", "TSAWWASSEN", "`WHITE ROCK`", "LANGLEY","`MAPLE RIDGE`","`BOWEN ISLAND`","`LIONS BAY`")

DF<-subset(df_dataYear, grepl(van, DEATHPC_6_CT))
DF$CITY<-trimws(DF$CITY)
DF2<- DF[DF$CITY %in% cities,] #236609-all data with death CT infor AND city information within Metro Van 

DF3<-subset(df_dataYear, grepl(van, PC_CT)) 
DF3$CITY<-trimws(DF3$CITY)
deathPCCTs <-transform(DF3, CT = ifelse(!is.na(PC_CT), PC_CT, 0.00))#350982- all data with residence CT in Metro Van

DF4<- DF3[DF3$CITY %in% cities,] #266789 - data with residence CT info in Metro Van AND city information

df_dataYear$CITY<-trimws(df_dataYear$CITY)
DF5<- df_dataYear[df_dataYear$CITY %in% cities,] #267508-all data with city information in cities of Metro Van
DF8<-setdiff(x=DF5, y=DF4) #719 deaths with city info but no PC_PCT info
DF9<-setdiff(x=DF8, y=DF2) ######262 Deaths without any PC_CT or DEATH_PC_CT but have right city


DF6<-setdiff(x=DF2, y=DF3) #457-Right DeathCTs and Cities but not PC_CTs
missDeathCTs <-transform(DF6, CT = ifelse(!is.na(DEATHPC_6_CT), DEATHPC_6_CT, 0.00)) #457 death files without PC_CT but does have right city - so we include


DF10<-setdiff(x=DF, y=DF3) #14674 all deaths with death PC info but no or incorrect city or residence info
DF10$CITY<-trimws(DF10$CITY)
DF10_blank<-subset(DF10, CITY=="")
DF10_unknown<-subset(DF10, CITY=="UNKNOWN" | CITY=="UNKNOWN, UNKNOWN")

DF_unknowns<-rbind(DF10_blank, DF10_unknown) #78 CTs with deaths in Van but no city info
deathPCretrieved <-transform(DF_unknowns, CT = ifelse(!is.na(DEATHPC_6_CT), DEATHPC_6_CT, 0.00))

deathsCityNoNA<-subset(DF5,!grepl(van,PC_CT) & !grepl(van,DEATHPC_6_CT)) 
deathsCityNoNA <-transform(deathsCityNoNA, CT = ifelse(!is.na(PC_CT), PC_CT, 0.00))
deathsCityNoNA<-subset(deathsCityNoNA, CT==0 | CT==590000000 | CT>599999999) #111 deaths with city info but incorrect PC_CT and death_CT - either NA or really high number
deathsCityNoNA$CT<-str_sub(deathsCityNoNA$CT,6)
deathsCityNoNA <-transform(deathsCityNoNA, CT = 600.00)



DF_final<-rbind(deathPCCTs, missDeathCTs, deathPCretrieved)###Merged all data with residence info, death info in Van but no correct city info,and death files without residence but does have right city and death cT



DF_final$CT<-str_sub(DF_final$CT,6)
DF_final$CT<-as.numeric(as.character(DF_final$CT))
all_data<-DF_final #351587 final death count 

```


## Include only CTs and variables of interest and save file

```{r}
deaths=all_data[all_data$CT %in% CT2011,]
finalMortality <-select(deaths,CT,PC_CT,DEATHPC_6_CT,DEATHDTCCYY,DEATHDTMM,SEX,MARITAL,NATIVE,BIRTHDTCCYY,AGE,BIRTHCITY,OCCUPATION,ICD_VERSN,CAUSE,AXIS_CODES,STUDYID)

write.csv(finalMortality, file="R:/working/preparing data/DeathData/final_mv_allyears.csv")
```

