---
title: "ImputeMSD"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                                         
  Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                    
                                                                                                         
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries and import file                                          
             (B) Functions for imputing and finding means                       
             (C) Merge material and social deprivation data                                     
             (D) Subset individual years of data
             (E) Finalise the MSD dataset and save file

#load libraries and import file        
             
```{r}
library(spdep)
library(dplyr)
msd<-read.csv("R:/working/preparing data/CovarData/msd_all.csv")

```

#functions for imputing and finding means

```{r}
impute<- function(data, score, yearchoice) {
  load("R:/working/modeling/shape_files/shape_478CTs.RData")
  load("R:/working/preparing data/CTIDS.RData")
  names(CTIDs)<-c("area","CT")
  CTIDs$CT<-as.numeric(as.character(CTIDs$CT))


  NB <- poly2nb(shape, queen=TRUE, row.names=shape$CTNAME)

  subsetList<-function(NB, elementNames) {
    lapply(elementNames, FUN=function(x) CTIDs$CT[NB[[x]]])
  }
  
  msd<-data %>%
    select(CT,year,score,population) %>%
    filter(year==yearchoice)
  msdCT<-select(msd, CT)
  
  
  missingMsdCT<-anti_join(CTIDs,msdCT, by="CT")
  #names(missingMsdCT)<-c("CT","area")
  missingMsdCT<-select(missingMsdCT, "area")
  
  for (i in missingMsdCT){ #include msdi CTs we want to merge
  NB2<-subsetList(NB, c(i))
  }
  
  msd<-msd %>%
    select(CT,score) %>%
    rename(x=CT)
  
  NB3<-lapply(NB2, function(x) merge(x, msd, by="x"))
  new.names<-c("CT",score)
  NB4<-lapply(NB3, setNames, new.names)  
  return(NB4)
}  



findMean<-function(data2, data,yearchoice, score) {
  load("R:/working/modeling/shape_files/shape_478CTs.RData")
  load("R:/working/preparing data/CTIDS.RData")
  names(CTIDs)<-c("area","CT")
  CTIDs$CT<-as.numeric(as.character(CTIDs$CT))


  NB <- poly2nb(shape, queen=TRUE, row.names=shape$CTNAME)

  subsetList<-function(NB, elementNames) {
    lapply(elementNames, FUN=function(x) CTIDs$CT[NB[[x]]])
  }
  
  msd<-data2 %>%
    select(CT,year,score,population) %>%
    filter(year==yearchoice)
  msdCT<-select(msd, CT)
  
  
  missingMsdCT<-anti_join(CTIDs,msdCT, by="CT")
  #names(missingMsdCT)<-c("CT","area")
  missingMsdCT<-select(missingMsdCT, "area")
  
  NB4<-data
  NB6<-lapply(NB4, function(x) x[1,])

  NB7<-lapply(NB6, function(x) x["mean"])
  print(NB7)
  Scores<-do.call(rbind.data.frame, NB7)
  names(Scores)<-score
  
  msdmeans<-missingMsdCT %>%
    cbind(Scores) %>%
    left_join(CTIDs, by="area") %>%
    select(CT, score)
  
  return(msdmeans)
 
}

```

#Merge material and social deprivation data  

```{r}

mergeMSD<-function(data, years) {

  msd<-data
  
  msd.soc<-impute(msd,
                      score = "SocScore",
                      yearchoice = years)
  
  msd.soc<-lapply(msd.soc, function(x) transform(x, mean=mean(SocScore)))
  
  msd.soc<-findMean(data2=msd,
                      data = msd.soc,
                      score = "SocScore",
                      yearchoice = years)
  
  msd.mat<-impute(msd,
                      score = "MatScore",
                      yearchoice = years)
  
  msd.mat<-lapply(msd.mat, function(x) transform(x, mean=mean(MatScore)))
  
  msd.mat<-findMean(data2=msd,
                      data = msd.mat,
                      score = "MatScore",
                      yearchoice = years)
  
  msdMiss<-merge(msd.mat, msd.soc, by="CT")
  
  msdMiss$year<-rep(years)
  
  load("R:/working/preparing data/PopulationData/pop_interpolated.RDATA")
  pops<-pop %>% filter(year==years) %>% group_by(CT) %>% mutate(population=sum(pop)) %>% select(CT,population) %>% unique()
  
  msdMiss<-msdMiss %>% left_join(pops, by="CT") %>% select(CT, year, MatScore, SocScore, population)
  
  return(msd<-msd %>% filter(year==years) %>% rbind(msdMiss))
}


```

#Subset individual years of data

```{r}
rm(merge)

msd1991<-mergeMSD(msd,
                  years = 1991)

msd1996<-mergeMSD(msd,
                  years = 1996)

msd2001<-mergeMSD(msd,
                  years = 2001)

msd2006<-mergeMSD(msd,
                  years = 2006)

msd2011<-mergeMSD(msd,
                  years = 2011)

msd2016<-mergeMSD(msd,
                  years = 2016)

msd<-rbind(msd1991,msd1996,msd2001,msd2006,msd2011,msd2016)

save(msd, file="R:/working/preparing data/CovarData/msd_processed.RDATA")
```

#Finalise the MSD dataset and save file

```{r}
load("R:/working/preparing data/CovarData/msd_processed.RDATA")
merge<-read.csv("R:/working/preparing data/Merge/merge_all12.csv", header=T)

library(car)
for (i in 1:nrow(merge)){
  msd$CT[which(msd$CT==as.numeric(merge[i,1]))]<-as.numeric(merge[i,2])
}
detach("package:car", unload=TRUE)
library(dplyr)

newMSD <- msd %>%
  group_by(CT,year) %>%
  summarise(SocScore=mean(SocScore, na.rm=TRUE), MatScore=mean(MatScore, na.rm=TRUE)) 
newMSD<-subset(newMSD,CT!=270)


msd<-newMSD[order(newMSD$CT),]


msd<-na.omit(msd)


length(unique(msd$CT))

save(msd, file="R:/working/preparing data/CovarData/msdcwed.RDATA")

```



