---
title: "prepareCovar"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                                         
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
                                                                                                         
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load all libraries and data                                          
             (B) Interpolate missing data using multiple years of data                        
             (C) Load and prepare crosswalk files                                               
             (D) Linearly interpolate data so all years are available
             (E) Finalise the covariate dataset and save file

## Load all libraries and data 

```{r}
library("data.table")
library("plyr")
library("dplyr")
library("tidyr")


#Create import function for any covariables

import<-function(covariable) {
  dir<-"R:/working/preparing data/CovarData/allyears/"  
  
  for (i in c(2016,2011,2006,2001,1996,1991)) {
    assign(paste0("covar", i), read.csv(paste0(dir, i,"/", covariable, i, ".csv"), header=T, sep=","))
    }
    
  covar1991$year <-rep(1991,nrow(covar1991))
  covar1996$year <-rep(1996,nrow(covar1996))
  covar2001$year <-rep(2001,nrow(covar2001))
  covar2006$year <-rep(2006,nrow(covar2006))
  covar2011$year <-rep(2011,nrow(covar2011))
  covar2016$year <-rep(2016,nrow(covar2016))

  covarfiles<-rbind(covar1991,covar1996,covar2001,covar2006,covar2011,covar2016)
  return(subset(covarfiles, CT>0))
}

#Import specific covariables and remove any years missing

education<-import('education')
education<-subset(education,year!=1996) #removed 1996 because it needs to be interpolated - can remove this later
education<-education %>%
  mutate(numerator=None+HS) %>% #creating variable that includes all below HS
  select(CT, year, numerator, Total)

aboriginal<-import('aboriginal')
aboriginal<-aboriginal %>%
#  subset(year!=2001) %>%#removed 2001 because it needs to be interpolated - can remove this later
  select(CT,year,Aboriginal,Population)


```

## Interpolate missing data using multiple years of data  

```{r}
generateMissing<-function(covariable,col,col2,name,name2) {
  covar270<-subset(covariable, CT==210 | CT==281.01 | CT==281.02 | CT==282.02 | CT==281.01| CT==283.00)#0
  covar280<-subset(covariable, CT==280.01)
  covar290<-subset(covariable, CT==290.06)
  covar250<-subset(covariable, CT==250.01) #only input for 2006 and 2016

#new imputations for missing covar files using historical data
  col270<-mean(covar270[[col]], na.rm=T)
  col280<-mean(covar280[[col]], na.rm = T)
  col290<-mean(covar290[[col]], na.rm = T)
  col250<-mean(covar250[[col]], na.rm = T)
  col270.2<-mean(covar270[[col2]], na.rm=T)
  col280.2<-mean(covar280[[col2]], na.rm = T)
  col290.2<-mean(covar290[[col2]], na.rm = T)
  col250.2<-mean(covar250[[col2]], na.rm = T)
  year270<-c(1991,1996,2001,2006,2011)
  CT270<-c(270)
  year280<-c(1996,2001)
  CT280<-c(280.01)
  year290<-c(1996)
  CT290<-c(290.06)
  year250<-c(2011)
  CT250<-c(250.01)
  missingCT270<-data.frame(CT270,year270,col270,col270.2)
  missingCT280<-data.frame(CT280,year280,col280,col280.2)
  missingCT290<-data.frame(CT290,year290,col290,col290.2)
  missingCT250<-data.frame(CT250,year250,col250,col250.2)
      
  names(missingCT270)<-c("CT","year", name, name2)
  names(missingCT280)<-c("CT","year", name, name2)
  names(missingCT290)<-c("CT","year", name, name2)
  names(missingCT250)<-c("CT","year", name, name2)

  return(missingCTs<-rbind(missingCT270, missingCT280, missingCT290, missingCT250))
}

#Remove missing files

missingEdu<-generateMissing(education, "numerator","Total", "HS", "population")
names(education)<-c("CT", "year", "HS", "population")
education<-rbind(education,missingEdu)
rm(missingEdu)


missingAb<-generateMissing(aboriginal, "Aboriginal","Population", "aboriginal", "population")
names(aboriginal)<-c("CT", "year", "aboriginal", "population")
aboriginal<-rbind(aboriginal,missingAb)
rm(missingAb)


```


#Load and prepare crosswalk files

```{r}
#load crosswalks

applycw<-function(covariable, col, col2) {

covariable<-aboriginal
col<-"aboriginal"
col2<-"population"
  
  crosswalk91<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk91.csv", header=TRUE)
  crosswalk96<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk96.csv", header=TRUE)
  crosswalk01<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk01.csv", header=TRUE)
  crosswalk06<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk06.csv", header=TRUE)
  crosswalk11<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk11.csv", header=TRUE)

  names(covariable)[1]<-"ctuid_s"
  
  for (i in c(2016,2011,2006,2001,1996,1991)) {
    assign(paste0("covar", i), subset(covariable, year==i))
    }
  names(covar2016)[1]<-"ctuid_t"


  
  covar91<-covar1991
  covar96<-covar1996
  covar01<-covar2001
  covar06<-covar2006
  covar11<-covar2011
  covar2016<-select(covar2016, "year","ctuid_t", col, col2)
  names(covar2016)<-c("year","ctuid_t","covarnew","covarnew2")
  
  
  #Multiply 1991 covar file with crosswalk table to get 2016 CT estimates
  covar91<-merge(x=crosswalk91, y=covar91, by="ctuid_s", all.y=TRUE)
    print(covar91)
  covar91<-subset(covar91, ctuid_s>0)
  covar91<-covar91[order(covar91$ctuid_s,covar91$year),]
  covar91[[col]] <- covar91[["w"]] * covar91[[col]] 
  covar91[[col2]] <- covar91[["w"]] * covar91[[col2]]
  covar1991b = covar91 %>% group_by(ctuid_t, year) %>%
    summarise_(covarnew  = paste0("sum(",col,", na.rm = T)"))
  covar1991c = covar91 %>% group_by(ctuid_t, year) %>%
  summarise_(covarnew2  = paste0("sum(",col2,", na.rm = T)"))
  covar1991<-merge(covar1991b,covar1991c, by="ctuid_t")
  names(covar1991)[2]<-"year"
  covar1991<-select(covar1991,"year","ctuid_t","covarnew","covarnew2")

  #Multiply 1996 covar file with crosswalk table to get 2016 CT estimates
  covar96<-merge(x=crosswalk96, y=covar96, by="ctuid_s", all.y=TRUE)
  covar96<-subset(covar96, ctuid_s>0)
  covar96<-covar96[order(covar96$ctuid_s,covar96$year),]
  covar96[[col]] <- covar96[["w"]] * covar96[[col]] 
  covar96[[col2]] <- covar96[["w"]] * covar96[[col2]]
  covar1996b = covar96 %>% group_by(ctuid_t, year) %>%
    summarise_(covarnew  = paste0("sum(",col,", na.rm = T)"))
  covar1996c = covar96 %>% group_by(ctuid_t, year) %>%
  summarise_(covarnew2  = paste0("sum(",col2,", na.rm = T)"))
  covar1996<-merge(covar1996b,covar1996c, by="ctuid_t")
  names(covar1996)[2]<-"year"
  covar1996<-select(covar1996,"year","ctuid_t","covarnew","covarnew2")
  
  #Multiply 2001 covar file with crosswalk table to get 2016 CT estimates
  covar01<-merge(x=crosswalk01, y=covar01, by="ctuid_s", all.y=TRUE)
  print(covar01)
  covar01<-subset(covar01, ctuid_s>0)
  covar01<-covar01[order(covar01$ctuid_s,covar01$year),]
  covar01[[col]] <- covar01[["w"]] * covar01[[col]] 
  covar01[[col2]] <- covar01[["w"]] * covar01[[col2]]
  covar2001b = covar01 %>% group_by(ctuid_t, year) %>%
    summarise_(covarnew  = paste0("sum(",col,", na.rm = T)"))
  covar2001c = covar01 %>% group_by(ctuid_t, year) %>%
  summarise_(covarnew2  = paste0("sum(",col2,", na.rm = T)"))
  covar2001<-merge(covar2001b,covar2001c, by="ctuid_t")
  names(covar2001)[2]<-"year"
  covar2001<-select(covar2001,"year","ctuid_t","covarnew","covarnew2")
  
  #Multiply 2006 covar file with crosswalk table to get 2016 CT estimates
  covar06<-merge(x=crosswalk06, y=covar06, by="ctuid_s", all.y=TRUE)
  covar06<-subset(covar06, ctuid_s>0)
  covar06<-covar06[order(covar06$ctuid_s,covar06$year),]
  covar06[[col]] <- covar06[["w"]] * covar06[[col]] 
  covar06[[col2]] <- covar06[["w"]] * covar06[[col2]]
  covar2006b = covar06 %>% group_by(ctuid_t, year) %>%
    summarise_(covarnew  = paste0("sum(",col,", na.rm = T)"))
  covar2006c = covar06 %>% group_by(ctuid_t, year) %>%
  summarise_(covarnew2  = paste0("sum(",col2,", na.rm = T)"))
  covar2006<-merge(covar2006b,covar2006c, by="ctuid_t")
  names(covar2006)[2]<-"year"
  covar2006<-select(covar2006,"year","ctuid_t","covarnew","covarnew2")

  
  #Multiply 2011 covar file with crosswalk table to get 2016 CT estimates
  covar11<-merge(x=crosswalk11, y=covar11, by="ctuid_s", all.y=TRUE)
  covar11<-subset(covar11, ctuid_s>0)
  covar11<-covar11[order(covar11$ctuid_s,covar11$year),]
  covar11[[col]] <- covar11[["w"]] * covar11[[col]]
  covar11[[col2]] <- covar11[["w"]] * covar11[[col2]]
  covar2011b = covar11 %>% group_by(ctuid_t, year) %>%
    summarise_(covarnew  = paste0("sum(",col,", na.rm = T)"))
  covar2011c = covar11 %>% group_by(ctuid_t, year) %>%
  summarise_(covarnew2  = paste0("sum(",col2,", na.rm = T)"))
  covar2011<-merge(covar2011b,covar2011c, by="ctuid_t")
  names(covar2011)[2]<-"year"
  covar2011<-select(covar2011,"year","ctuid_t","covarnew","covarnew2")

  
  covarcwed<-rbind(covar1991,covar1996,covar2001,covar2006,covar2011,covar2016)
  names(covarcwed)<-c("year","CT",col,col2)
  

  
  return(covarcwed)
  }

#Apply crosswalk to covariable files and create education variable

aboriginalcwed<-covarcwed

aboriginalcwed<-applycw(aboriginal, "aboriginal","population")

save(aboriginalcwed, file="R:/working/preparing data/CovarData/aboriginalcwed.RDATA")


```

#Merge Certain CTs

```{r}
library(dplyr)
load("R:/working/preparing data/CovarData/aboriginalcwed.RDATA")

merge<-function(covariable, col, col2) {
  load(file="R:/working/preparing data/CTIDs.RData")
  merge<-read.csv("R:/working/preparing data/Merge/merge_all12.csv", header=T)
  
  for (i in 1:nrow(merge)){
    covariable$CT[which(covariable$CT==as.numeric(merge[i,1]))]<-as.numeric(merge[i,2])
  }
  
  covar <-covariable %>%
    group_by(CT,year) %>%
    summarise_if(is.numeric, sum, na.rm=TRUE)
  
  covar<-subset(covar,CT!=270)
  names(covar)<-c("CT","year",col,col2)
  
  return(covar)
}

#apply merge to covariables

educationM<-merge(educationcwed, 'HS', 'population')
educationM<-educationM %>%
  mutate(education = HS/population) %>%
  select(year,CT,education)

aboriginalM<-merge(aboriginalcwed, 'aboriginal', 'population')
aboriginalM<-aboriginalM %>%
  mutate(Aboriginal = aboriginal/population) %>%
  select(year,CT,Aboriginal)

length(unique(aboriginalM$CT))

```


#Linearly interpolate data so all years are available

```{r}
load("R:/working/preparing data/CovarData/msdcwed.RDATA")

#create a function to linearly interpolate data where 2001 data is unavailable
interpolate2<-function(data, col) {
  
  covariable<-as.data.frame(data)
  
  covar91<-subset(covariable,year==1991)
  covar96<-subset(covariable,year==1996)
  covar01<-subset(covariable,year==2001)
  covar06<-subset(covariable,year==2006)
  covar11<-subset(covariable,year==2011)
  covar16<-subset(covariable,year==2016)


  
  covar91<-select(covar91,"CT",col)
  covar96<-select(covar96,"CT",col)
  covar01<-select(covar01,"CT",col)
  covar06<-select(covar06,"CT",col)
  covar11<-select(covar11,"CT",col)
  covar16<-select(covar16,"CT",col)
  

  
  names(covar91)<-c("CT","covar91")
  names(covar96)<-c("CT","covar96")
  names(covar01)<-c("CT","covar01")
  names(covar06)<-c("CT","covar06")
  names(covar11)<-c("CT","covar11")
  names(covar16)<-c("CT","covar16")
  
  #1991-1996
  covar9691<-cbind(covar96, covar91)
  
  covar9691<-covar9691[,c(1,2,4)]
  diff9691<-(covar9691$covar96-covar9691$covar91)/5
  
  covar9691$covar90<-covar9691$covar91#just replicated 1990 interpolations because negative numbers
  covar9691$covar92<-covar9691$covar91+diff9691
  covar9691$covar93<-covar9691$covar92+diff9691
  covar9691$covar94<-covar9691$covar93+diff9691
  covar9691$covar95<-covar9691$covar94+diff9691
  
  names(covar9691)[1]<-"CT"
  
  covar90<-covar9691[,c("CT","covar90")]
  covar92<-covar9691[,c("CT","covar92")]
  covar93<-covar9691[,c("CT","covar93")]
  covar94<-covar9691[,c("CT","covar94")]
  covar95<-covar9691[,c("CT","covar95")]
  
  #1996-2001
  covar0196<-cbind(covar06,covar96)
  covar0196<-covar0196[,c(1,2,4)]
  diff0196<-(covar0196$covar06-covar0196$covar96)/5
  
  covar0196$covar97<-covar0196$covar96+diff0196
  covar0196$covar98<-covar0196$covar97+diff0196
  covar0196$covar99<-covar0196$covar98+diff0196
  covar0196$covar00<-covar0196$covar99+diff0196
  
  names(covar0196)[1]<-"CT"
  
  covar97<-covar0196[,c("CT","covar97")]
  covar98<-covar0196[,c("CT","covar98")]
  covar99<-covar0196[,c("CT","covar99")]
  covar00<-covar0196[,c("CT","covar00")]

  #2001-2006
  covar0601<-cbind(covar06,covar01)
  covar0601<-covar0601[,c(1,2,4)]
  diff0601<-(covar0601$covar06-covar0601$covar01)/5
  
  covar0601$covar02<-covar0601$covar01+diff0601
  covar0601$covar03<-covar0601$covar02+diff0601
  covar0601$covar04<-covar0601$covar03+diff0601
  covar0601$covar05<-covar0601$covar04+diff0601
  
  names(covar0601)[1]<-"CT"
  
  covar02<-covar0601[,c("CT","covar02")]
  covar03<-covar0601[,c("CT","covar03")]
  covar04<-covar0601[,c("CT","covar04")]
  covar05<-covar0601[,c("CT","covar05")]
  
  #2006-2011
  covar0611<-cbind(covar06,covar11)
  covar0611<-covar0611[,c(1,2,4)]
  diff1106<-(covar0611$covar11-covar0611$covar06)/5
  
  covar0611$covar07<-covar0611$covar06+diff1106
  covar0611$covar08<-covar0611$covar07+diff1106
  covar0611$covar09<-covar0611$covar08+diff1106
  covar0611$covar10<-covar0611$covar09+diff1106
  
  names(covar0611)[1]<-"CT"
  
  covar07<-covar0611[,c("CT","covar07")]
  covar08<-covar0611[,c("CT","covar08")]
  covar09<-covar0611[,c("CT","covar09")]
  covar10<-covar0611[,c("CT","covar10")]
  
  #2011-2016
  covar1116<-cbind(covar16,covar11)
  covar1116<-covar1116[,c(1,2,4)]
  diff1611<-(covar1116$covar16-covar1116$covar11)/5
  
  covar1116$covar12<-covar1116$covar11+diff1611
  covar1116$covar13<-covar1116$covar12+diff1611
  covar1116$covar14<-covar1116$covar13+diff1611
  covar1116$covar15<-covar1116$covar14+diff1611
  
  names(covar1116)[1]<-"CT"
  
  covar12<-covar1116[,c("CT","covar12")]
  covar13<-covar1116[,c("CT","covar13")]
  covar14<-covar1116[,c("CT","covar14")]
  covar15<-covar1116[,c("CT","covar15")]


  
  covar<-list(covar00,covar01,covar02,covar03,covar04,covar05,covar06,covar07,covar08,covar09,covar10,covar11,covar12,covar13,covar14,covar15,covar16,covar90,covar91,covar92,covar93,covar94,covar95,covar96,covar97,covar98,covar99)

  
  colnames<-c("CT", col)
  covar<-lapply(covar, setNames, colnames)
  years<-list(2000,2001,2002,2003,2004,2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,1990,1991,1992,1993,1994,1995,1996,1997,1998,1999)
  
  for (i in c(1:27)) {
    covar[[i]]["year"]<-years[[i]]
  }
  
  covar<-plyr::ldply(covar, data.frame)
  
  covar<-covar %>% select(year,CT,col) %>% arrange(year,CT)

  return(covar)
}

aboriginalInt<-interpolate2(aboriginalM, 'Aboriginal')

rm(merge)
msd.soc<-interpolate2(msd, 'SocScore')
msd.mat<-interpolate2(msd, 'MatScore')
msd<-merge(msd.soc, msd.mat, by=c("year","CT"))
msd<-msd %>% arrange(year,CT)

save(aboriginalInt, file="R:/working/preparing data/CovarData/aboriginalInt.RDATA")
save(msd, file="R:/working/preparing data/CovarData/msdInt.RDATA")

```



#Finalise the covariate dataset and save file

```{r}
load("R:/working/preparing data/CovarData/aboriginalInt.RDATA")
load("R:/working/preparing data/CovarData/msdInt.RDATA")
covar<-left_join(aboriginalInt,msd, by=c("year","CT")) 
stdize = function(x,...){(x - min(x, ...)) / (max(x, ...) - min(x, ...))}
covar <- covar %>%
  mutate(msd=SocScore+MatScore) 
covar$msd<-stdize(covar$msd, na.rm=T)


covar$SocScore<-1-covar$SocScore
covar$MatScore<-1-covar$MatScore
covar$msd<-1-covar$msd

covar<-covar %>%
  group_by(CT)
covar$CT<-as.numeric(as.factor(covar$CT))
covar$CT<-(covar$CT-1)
covar$CT<-as.integer(covar$CT)
covar<-data.table::setDT(covar)
save(covar, file="R:/working/modeling/cov_files/covar_368CTs.RData")
```

