---
title: "AdjacencyMatrix"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca 
               
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium 
                     
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries and import data                                        
             (B) Create adjacency matrix with contiguous neighbours   
             (C) Plot adjacency matrix                                                                      

## Load libraries and import data

```{r}
library("spdep")
library("rgdal")
library("Matrix")
load("R:/working/modeling/shape_files/shape-mapping368CTs.RData")

```

## Create adjacency matrix with contiguous neighbours

```{r}
MV<-shape
NB <- poly2nb(MV, queen=TRUE, row.names=MV$FIPSNO)
#find any CTs that don't have neighbours
noNeigh <- which(card(NB)==0)
#find closest neighbour to any that has no bordering:
nearNeigh <- knn2nb(knearneigh(coordinates(MV), k=1))


CTs2016 <- nb2mat(NB, style="B", zero.policy=FALSE)
#make Matrix symmetric
CTs2016[lower.tri(CTs2016)] = t(CTs2016)[lower.tri(CTs2016)]
#make rownames and colnames symmetric
x <-CTs2016
colnames(x)<-rownames(x)
adjmat1<-Matrix(x, sparse=TRUE)
adjmat<-as(adjmat1, "dgTMatrix")
save(adjmat, file="R:/working/preparing data/AdjacencyMatrix/adjmat2016_368CTs.RData")
```


## Plot adjacency matrix

```{r}
library(igraph)

adj<-graph.adjacency(adjmatrix = adjmat1, mode="undirected", weighted=TRUE, diag=FALSE)

pdf("R:/working/preparing data/AdjacencyMatrix/adjmatrix.pdf", height=14, width=14)
plot.igraph(adj, vertex.size=2, vertex.label.cex=.5)
dev.off()
```

