---
title: "adjustDeathCTs"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import files, and define attributes                                                      
             (B) Adjust CTs of death causes and then group and save for all causes                                                               


## Load libraries, import files, and define attributes

```{r}

libs<-c("car","data.table","dplyr")
lapply(libs, require, character.only=TRUE)

merge<-read.csv("R:/working/preparing data/Merge/merge_all12.csv", header=T)

load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)

dir<-"R:/working/preparing data/DeathData/causes/deathfiles/crosswalked/"
file<-"_deaths.RData"
dir2<-"R:/working/modeling/deaths/causes/"

```


## Adjust CTs of death causes and then group and save files for all causes

```{r}

#Rename CT areas so that we can end up with df with area named as CT
load("R:/working/preparing data/CT368.RDATA")
names(CT)<-c("CT","area")

for (x in 1:21) {
  load(paste0(dir, causenames[x], file)) 
  
  for (i in 1:nrow(merge)){
    deaths$CT[which(deaths$CT==as.numeric(merge[i,1]))]<-as.numeric(merge[i,2])
  }

  deaths<-deaths %>%
    rename(area=CT) 
  
  deaths<-merge(deaths,CT, by="area")
  
  deaths<-deaths %>%
    select(CT,year,sex,age,death)
  
  newDeaths <-deaths %>%
    group_by(CT,year,sex,age) %>%
    summarise(death=sum(death, na.rm=TRUE))
  
  deaths<-newDeaths[order(newDeaths$CT),]
  
  deaths$CT<-as.numeric(deaths$CT)
  deaths$sex<-as.integer(deaths$sex)
  deaths<-na.omit(deaths)
  deaths<-data.table::setDT(deaths)
  
  length<-length(unique(deaths$CT))
  
  save(deaths, file=paste0(dir2, causenames[x], file))
}

```
