---
title: "categoriseAllDeaths"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                   
             (B) Loop the application of crosswalk files

## Load libraries, import data, and define attributes

```{r}
library("dplyr")
library("tidyr")
library("data.table")

rm(list=ls())
load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)

dir<-"R:/working/preparing data/DeathData/causes/deathfiles/"
file<-"_deaths.RData"
dir2<-"R:/working/preparing data/DeathData/causes/deathfiles/crosswalked/"
```

## Loop the application of crosswalk files

```{r}

for (i in 1:21) {
  load(paste0(dir, causenames[i], file)) 
  deaths <- deaths %>%
    filter(year>1980) %>%
    rename(ctuid_s=CT, death=deaths) 
  
  #import 2011 crosswalk file
  crosswalk11<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk11.csv", header=TRUE)

  #Multiply 2011 death file with crosswalk table to obtain 2016 CT estimates
  death2011<-merge(x=crosswalk11, y=deaths, by="ctuid_s", all.y=TRUE)
  death2016 = death2011 %>% 
    filter(ctuid_s>0) %>%
    arrange(ctuid_s, year, sex, age) %>%
    mutate(death2011 = w * death) 
  deaths<-death2016 %>%
    group_by(ctuid_t, year, sex, age) %>%
    summarise(death2011 = sum(death2011, na.rm = T)) %>%
    rename(CT=ctuid_t,death=death2011) %>%
    arrange(year,CT,sex,age,death)
  
  save(deaths, file=paste0(dir2, causenames[17], file))
}

```

