---
title: "categoriseAllDeaths"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Categorise age groups
             (C) Keep final columns and change variable types
             (D) Derive causes for deaths using ICD10 codes
             (E) Count number of deaths and save file
             
## Load libraries, import data, and define attributes  

```{r}
library("dplyr")
library("tidyr")
library("data.table")

setwd("R:/working/preparing data/DeathData")
allDeaths<-read.csv(file="final_mv_allyears.csv", header=TRUE, check.names = FALSE)
```


## Categorise age groups

```{r}
#Categorise age groups and rename column names
agebreaks <-c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,113)
agelabels <-c("0","5","10","15","20","25","30","35","40","45","50","55","60","65","70","75","80","85")

deaths<-setDT(allDeaths) [, agegroups := cut(AGE, breaks = agebreaks, right= FALSE, labels = agelabels)]

```

## Keep final columns and change variable types

```{r}

colnames(deaths)[colnames(deaths)=="DEATHDTCCYY"]<-"year"
colnames(deaths)[colnames(deaths)=="SEX"]<-"sex"
colnames(deaths)[colnames(deaths)=="agegroups"]<-"age"
colnames(deaths)[colnames(deaths)=="CAUSE"]<-"cause"
colnames(deaths)[colnames(deaths)=="ICD_VERSN"]<-"icdversn"



#Rename sex variables from factors to values
levels(deaths$sex)<- c("2","1","3")

deaths$sex<-as.numeric(as.character(deaths$sex))
deaths$age<-as.numeric(as.character(deaths$age))

deaths <-select(deaths,year,CT,sex,age,cause,icdversn)

```


## Derive causes for deaths using ICD10 codes

```{r}
load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(list.files(pattern="*.csv"),1,nchar(causefiles)-4)
load("R:/working/preparing data/DeathData/icd9/gbdCausesicd9.RData")
load("R:/working/preparing data/DeathData/gbdcausesicd10.RData")


#ICD10 codes
deaths_icd10<-subset(deaths, icdversn==10)
names(deaths_icd10)[5]<-"ICD10"
deaths_icd10$causeid<-0
deaths_icd10$causeidgarb<-0

deaths_icd10$ICD10 <- trimws(as.character(deaths_icd10$ICD10))
for (y in causenames) {
  deaths_icd10$causeid[which(deaths_icd10$ICD10 %in% gbdCausesicd10[[y]])] <- y
}

deathsbyCause10<-deaths_icd10 %>%
  group_by(causeid) %>%
  tally()

save(deaths_icd10, file="R:/working/preparing data/DeathData/causes/deaths_icd10.RData")


#ICD9 codes
deaths_icd9<-subset(deaths, icdversn==9)
names(deaths_icd9)[5]<-"ICD9"
deaths_icd9$causeid<-0

#trial with just three characters
deaths_icd9$ICD9<-as.character(deaths_icd9$ICD9)
deaths_icd9$ICD9<-gsub("(^\\s+)|(\\s+$)", "", deaths_icd9$ICD9)

for (y in causenames) {
  deaths_icd9$causeid[which(deaths_icd9$ICD9 %in% gbdCausesicd9[[y]])] <- y
}

deathsbyCause9<-deaths_icd9 %>%
  group_by(causeid) %>%
  tally()




for (y in causenames) {
temp<-transform(deaths_icd9,causeid=grepl(paste0(gbdCausesicd9[[y]], collapse = "|"),ICD9)+0)
deaths_icd9$causeid[temp$causeid == 1] <- y
}

deathsbyCause9<-deaths_icd9 %>%
  group_by(causeid) %>%
  tally()

save(deaths_icd9, file="R:/working/preparing data/DeathData/causes/deaths_icd9.RData")


```


## Count number of deaths and save file

```{r}
#Count number of deaths
mortcounts<-count(deaths, year, CT, sex, age)
colnames(mortcounts)[colnames(mortcounts)=="n"]<-"death"

#sort data by sequentially of CT
deaths<-mortcounts[order(mortcounts$CT),]

setDT(deaths)
write.csv(deaths, file="R:/working/preparing data/DeathData/years/allyears/allDeaths.csv")
save(deaths, file="R:/working/preparing data/DeathData/years/final/deaths.RData")

```
