---
title: "CategoriseICD9"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Population Data BC, BC Ministry of Health: https://www.popdata.bc.ca/                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                           
             (B) Assess patterns of ICD codes and categorise death files back into GBD cause groups           
             (C) Create ICD9 list                                                                      


## Load libraries, import data, and define attributes 

```{r}
library(icd)
library(tidyverse)
library(stringr)
load("R:/working/preparing data/DeathData/causes/deaths_icd9.Rdata")
load("R:/working/preparing data/DeathData/gbdcausesicd9.RData")
load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)
```

## Assess patterns of ICD codes and categorise death files back into GBD cause groups

```{r}
#trial with just max four characters
deaths_icd9$ICD9<-as.character(deaths_icd9$ICD9)
deaths_icd9$ICD9<-gsub("(^\\s+)|(\\s+$)", "", deaths_icd9$ICD9)

trial<-sapply(gbdCausesicd9, function(x) ifelse(nchar(x)>3, substr(x,1,nchar(x)-1), x))
deaths_icd9$code<-0
deaths_icd9$code<-ifelse(nchar(deaths_icd9$ICD9)>3, substr(deaths_icd9$ICD9,1,nchar(deaths_icd9$ICD9)-1), deaths_icd9$ICD9)

#remove unnecessary columns
deaths_icd9<-subset(deaths_icd9, select = -c(icdversn))

deaths_icd9$newcause<-0
for (y in causenames) {
  deaths_icd9$newcause[which(deaths_icd9$code %in% trial[[y]])] <- y
}

deaths_icd9$trial<-0
deaths_icd9$trial<-ifelse(deaths_icd9$causeid==0, deaths_icd9$newcause, deaths_icd9$causeid)

#Tally up
deathsbyCause9<-deaths_icd9 %>%
  group_by(trial) %>%
  tally()


#Next trial with just max three characters
trial2<-sapply(gbdCausesicd9, function(x) ifelse(nchar(x)>4, substr(x,1,nchar(x)-2), x))
trial2<-sapply(trial2, function(x) ifelse(nchar(x)>3, substr(x,1,nchar(x)-1), x))

trial2<-sapply(trial2, function(x) ifelse(nchar(x)<2,NA, x))
deaths_icd9$code2<-ifelse(nchar(deaths_icd9$ICD9)>4, substr(deaths_icd9$ICD9,1,nchar(deaths_icd9$ICD9)-2), deaths_icd9$ICD9)
deaths_icd9$code2<-ifelse(nchar(deaths_icd9$code2)>3, substr(deaths_icd9$code2,1,nchar(deaths_icd9$code2)-1), deaths_icd9$code2)

deaths_icd9$newcause2<-0
for (y in causenames) {
  deaths_icd9$newcause2[which(deaths_icd9$code2 %in% trial2[[y]])] <- y
}
deaths_icd9$trial2<-0
deaths_icd9$trial2<-ifelse(deaths_icd9$trial==0, deaths_icd9$newcause2, deaths_icd9$trial)

deathsbyCause9<-deaths_icd9 %>%
  group_by(trial2) %>%
  tally()

#Next trial with just max two characters
trial3<-sapply(trial2, function(x) ifelse(nchar(x)>2, substr(x,1,nchar(x)-1), x))

#remove certain codes that will be problematic because of duplicates in other disease lists
#MSK
duplicatesMSK<-c("41", "43", "44", "69")
trial3[["msk"]]<-trial3[["msk"]][!(trial3[["msk"]] %in% duplicatesMSK)]
#otherNCD
duplicatesNCD<-c("55", "56")
trial3[["otherncd"]]<-trial3[["otherncd"]][!(trial3[["otherncd"]] %in% duplicatesNCD)]
duplicatesDIGEST<-c("78")
trial3[["digest"]]<-trial3[["digest"]][!(trial3[["digest"]] %in% duplicatesDIGEST)]
duplicatesDIAB<-c("77")
trial3[["diab_ckd"]]<-trial3[["diab_ckd"]][!(trial3[["diab_ckd"]] %in% duplicatesDIAB)]
duplicatesHIV<-c("54")
trial3[["hiv_std"]]<-trial3[["hiv_std"]][!(trial3[["hiv_std"]] %in% duplicatesHIV)]
duplicatesNEO<-c("56")
trial3[["neo"]]<-trial3[["neo"]][!(trial3[["neo"]] %in% duplicatesNEO)]
duplicatesNEURO<-c("77")
trial3[["neuro"]]<-trial3[["neuro"]][!(trial3[["neuro"]] %in% duplicatesNEURO)]
duplicatesNTD<-c("42")
trial3[["ntd"]]<-trial3[["ntd"]][!(trial3[["ntd"]] %in% duplicatesNTD)]
duplicatesNUTRITION<-c("24","71")
trial3[["nutrition"]]<-trial3[["nutrition"]][!(trial3[["nutrition"]] %in% duplicatesNUTRITION)]
duplicatesRI<-c("34","79")
trial3[["respinfect"]]<-trial3[["respinfect"]][!(trial3[["respinfect"]] %in% duplicatesRI)]
duplicatesSKIN<-c("13","45")
trial3[["skin"]]<-trial3[["skin"]][!(trial3[["skin"]] %in% duplicatesSKIN)]
duplicatesSUBS<-c("35","76","79")
trial3[["skin"]]<-trial3[["skin"]][!(trial3[["skin"]] %in% duplicatesSUBS)]
duplicatesUNINTENT<-c("45")
trial3[["unintent"]]<-trial3[["unintent"]][!(trial3[["unintent"]] %in% duplicatesUNINTENT)]



deaths_icd9$code3<-ifelse(nchar(deaths_icd9$code2)>2, substr(deaths_icd9$code2,1,nchar(deaths_icd9$code2)-1), deaths_icd9$code2)


deaths_icd9$newcause3<-0
for (y in causenames) {
  deaths_icd9$newcause3[which(deaths_icd9$code3 %in% trial3[[y]])] <- y
}
deaths_icd9$causes<-0
deaths_icd9$causes<-ifelse(deaths_icd9$trial2==0, deaths_icd9$newcause3, deaths_icd9$trial2)

save(deaths_icd9, file="R:/working/preparing data/DeathData/garbagecodes/deaths_icd9.RDATA")

deathsbyCause9<-deaths_icd9 %>%
  group_by(causes) %>%
  tally()

deathsbyCauses9final<-select(deaths_icd9, year, CT, sex, age, causes)

#Isolate causes with no causes identified
noCauses<-subset(deaths_icd10, causes==0)
noCauses<-select(noCauses,1:5)
colnames(noCauses)[5]<-"code"
noCauses<-left_join(noCauses, icd10cm2019, by="code")

noCauseTally<-noCauses %>%
  group_by(major) %>%
  tally()

save(deathsbyCauses9final, file="R:/working/preparing data/DeathData/garbagecodes/icd9/deathsbyCausesICD9_final.RDATA")
save(noCauses, file="R:/working/preparing data/DeathData/garbagecodes/noCausesICD9.RDATA")
save(noCauseTally, file="R:/working/preparing data/DeathData/garbagecodes/noCauseICD9Tally.RDATA")

```

## Create ICD9 list

```{r}
deaths_icd9_list<-deaths_icd9 %>% select(causes, ICD9) %>% filter(causes!=0) %>% group_by(causes) %>% summarise(ICD9=paste(unique(ICD9), collapse=', '))

write.csv(deaths_icd9_list, "R:/working/preparing data/DeathData/garbagecodes/deaths_icd9_list.csv", row.names = F)

```


