---
title: "categoriseICD10"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Assess patterns of ICD codes and categorise death files back into GBD cause groups                                              

             
## Load libraries, import data, and define attributes        

```{r}
library(icd)
library(tidyverse)
load("R:/working/preparing data/DeathData/causes/deaths_icd10.Rdata")
load("R:/working/preparing data/DeathData/causefiles.RData")
load("R:/working/preparing data/DeathData/gbdcausesicd10.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)
```

## Assess patterns of ICD codes and categorise death files back into GBD cause groups  

```{r}
#trial with just three characters
trial<-sapply(gbdCausesicd10, function(x) ifelse(nchar(x)>3, substr(x,1,nchar(x)-1), x))
deaths_icd10$code<-ifelse(nchar(deaths_icd10$ICD10)>3, substr(deaths_icd10$ICD10,1,nchar(deaths_icd10$ICD10)-1), deaths_icd10$ICD10)

deaths_icd10<-subset(deaths_icd10, select = -c(icdversn, causeidgarb))

deaths_icd10$newcause<-0
for (y in causenames) {
  deaths_icd10$newcause[which(deaths_icd10$code %in% trial[[y]])] <- y
}

deaths_icd10$cause<-0
deaths_icd10$cause<-ifelse(deaths_icd10$causeid==0, deaths_icd10$newcause, deaths_icd10$causeid)

#Next trial with just two characters
trial2<-sapply(gbdCausesicd10, function(x) ifelse(nchar(x)>2, substr(x,1,nchar(x)-2), x))
trial2<-sapply(trial2, function(x) ifelse(nchar(x)<2,NA, x))
deaths_icd10$code2<-ifelse(nchar(deaths_icd10$ICD10)>3, substr(deaths_icd10$ICD10,1,nchar(deaths_icd10$ICD10)-2), deaths_icd10$ICD10)
deaths_icd10$code2<-ifelse(nchar(deaths_icd10$code2)>2, substr(deaths_icd10$code2,1,nchar(deaths_icd10$code2)-1), deaths_icd10$code2)

deaths_icd10$newcause2<-0
for (y in causenames) {
  deaths_icd10$newcause2[which(deaths_icd10$code2 %in% trial2[[y]])] <- y
}
deaths_icd10$causes<-0
deaths_icd10$causes<-ifelse(deaths_icd10$cause==0, deaths_icd10$newcause2, deaths_icd10$cause)

save(deaths_icd10, file="R:/working/preparing data/DeathData/garbagecodes/deaths_icd10.RDATA")

#Tally up
deathsbyCause10<-deaths_icd10 %>%
  group_by(newcause) %>%
  tally()

deathsbyCauseFinal<-deaths_icd10 %>%
  group_by(causes) %>%
  tally()

#Isolate causes with no causes identified
noCauses<-subset(deaths_icd10, causes==0)
noCauses<-select(noCauses,1:5)
colnames(noCauses)[5]<-"code"
noCauses<-left_join(noCauses, icd10cm2019, by="code")

noCauseTally<-noCauses %>%
  group_by(major) %>%
  tally()

deathsbyCauses10final<-select(deaths_icd10, year, CT, sex, age, causes)

save(deathsbyCauses10final, file="R:/working/preparing data/DeathData/garbagecodes/icd10/deathsbyCauseICD10.RDATA")
save(deathsbyCauseFinal, file="R:/working/preparing data/DeathData/garbagecodes/deathsbyCause10.RDATA")
save(noCauses, file="R:/working/preparing data/DeathData/garbagecodes/noCauses.RDATA")
save(noCauseTally, file="R:/working/preparing data/DeathData/garbagecodes/noCauseICD10Tally.RDATA")
```

