---
title: "SaveCauses"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Combine death files and remove death files without causes                                        
             (C) Group deaths by causes                                                                   
             (D) Link death files with gbd codes
             (E) Save individual cause-specific data frames

## Load libraries, import data, and define attributes  
             
```{r}
library(dplyr)
library(data.table)
load("R:/working/preparing data/DeathData/garbagecodes/icd10/deathsbyCauseICD10.RDATA")
load("R:/working/preparing data/DeathData/garbagecodes/icd9/deathsbyCausesICD9_final.RDATA")

load("R:/working/preparing data/DeathData/causefiles.RData")
causenames=substr(causefiles,1,nchar(causefiles)-4)

causeids<-read.csv(file="R:/working/preparing data/DeathData/causes/gbdcauseids.csv", header=T)
```

## Combine death files and remove death files without causes

```{r}
allCauseDeaths<-rbind(deathsbyCauses9final,deathsbyCauses10final)
allCauseDeaths<-subset(allCauseDeaths, causes!=0)

save(allCauseDeaths, file="R:/working/preparing data/DeathData/causes/allCauseDeaths.rdata")
```


## Group deaths by causes

```{r}
causecounts=paste0(substr(causefiles,1,nchar(causefiles)-4),"_deaths")

#Create individual dataframes for each cause
for (y in causenames) {
  assign(paste0(y,"_deaths"),subset(allCauseDeaths,causes==paste0(y)))
}

deathsByCause<-mget(ls(pattern='+_deaths'))
#deathsByCause2<-lapply(deathsByCause, function(x) { x[["causes"]]<-NULL; x })
#list2env(deathsByCause2,.GlobalEnv)

deathsByCause3<-lapply(deathsByCause, function(x) count(x, year, CT, sex, age,causes))
colnames<-c("year","CT","sex","age","acause","death")
deathsByCause3<-lapply(deathsByCause3, setNames,colnames)
deathsByCause3<-lapply(deathsByCause3, function(df){
                       df[order(df$CT),]
                       })
deathsByCause<-lapply(deathsByCause3, function(df) setDT(df))

allDeaths<-do.call(rbind.data.frame, deathsByCause)

```

## Link death files with gbd codes

```{r}
death_and_ids<-left_join(x=allDeaths, y=causeids, by="acause")
death_and_ids<-select(death_and_ids, year,CT,sex,age,level,parent_id,cause_id,acause,death,level_0,level_1,level_2)
deaths<-death_and_ids

#Save full deaths list with causes
save(deaths, file="R:/working/preparing data/DeathData/causes/all/deaths.Rdata")


```


## Save individual cause-specific data frames

```{r}
list2env(deathsByCause,.GlobalEnv)

#Save in CSV format
lapply(names(deathsByCause), function(X) write.csv(deathsByCause[[X]], file=paste0(X, ".csv")))

#Save in RDATA format
lapply(names(deathsByCause), function(X) {
  assign(X, deathsByCause[[X]])
  save(list=X, file=paste0(X, ".Rdata"))
})

#Rename all cause lists to deaths and save

dir<-"R:/working/preparing data/DeathData/causes/deathfiles/"

deaths<-cvd_deaths
save(deaths, file = paste0(dir,"cvd_deaths.RData"))
deaths<-msk_deaths
save(deaths, file = paste0(dir,"msk_deaths.RData"))
deaths<-skin_deaths
save(deaths, file = paste0(dir,"skin_deaths.RData"))
deaths<-diab_ckd_deaths
save(deaths, file = paste0(dir,"diab_ckd_deaths.RData"))
deaths<-digest_deaths
save(deaths, file = paste0(dir,"digest_deaths.RData"))
deaths<-enteric_all_deaths
save(deaths, file = paste0(dir,"enteric_all_deaths.RData"))
deaths<-hiv_std_deaths
save(deaths, file = paste0(dir,"hiv_std_deaths.RData"))
deaths<-infect_deaths
save(deaths, file = paste0(dir,"infect_deaths.RData"))
deaths<-inj_trans_deaths
save(deaths, file = paste0(dir,"inj_trans_deaths.RData"))
deaths<-intent_deaths
save(deaths, file = paste0(dir,"intent_deaths.RData"))
deaths<-mater_neonat_deaths
save(deaths, file = paste0(dir,"mater_neonat_deaths.RData"))
deaths<-mental_deaths
save(deaths, file = paste0(dir,"mental_deaths.RData"))
deaths<-neo_deaths
save(deaths, file = paste0(dir,"neo_deaths.RData"))
deaths<-neuro_deaths
save(deaths, file = paste0(dir,"neuro_deaths.RData"))
deaths<-ntd_deaths
save(deaths, file = paste0(dir,"ntd_deaths.RData"))
deaths<-nutrition_deaths
save(deaths, file = paste0(dir,"nutrition_deaths.RData"))
deaths<-otherncd_deaths
save(deaths, file = paste0(dir,"otherncd_deaths.RData"))
deaths<-resp_deaths
save(deaths, file = paste0(dir,"resp_deaths.RData"))
deaths<-respinfect_deaths
save(deaths, file = paste0(dir,"respinfect_deaths.RData"))
deaths<-subs_deaths
save(deaths, file = paste0(dir,"subs_deaths.RData"))
deaths<-unintent_deaths
save(deaths, file = paste0(dir,"unintent_deaths.RData"))

```


