---
title: "mergeShapefiles"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries and import data                                                    
             (B) Apply algorithm to replace old CTIDs with new CTIds from another dataframe

#Load libraries and import data

```{r}
libs<-c("rgdal", "maptools", "gridExtra", "raster", "sp", "dplyr", "rgeos")
lapply(libs, require, character.only=TRUE)

shape2016<-readOGR(dsn="R:/working/preparing data/ShapeFiles/allyears/2016_CensusTracts/Selection/2016MetroVanCTs.shp", layer="2016MetroVanCTs")
shape2016$CTNAME<-as.numeric(as.character(shape2016$CTNAME))

merge<-read.csv(file="R:/working/preparing data/Merge/merge_1.csv", header=T)
```

             
#Apply algorithm to replace old CTIDs with new CTIds from another dataframe

```{r}

outliersremove<-dplyr::select(merge,1)
outliersmerge<-dplyr::select(merge,2)
names(outliersremove)<-names(outliersmerge)
alloutliers<-rbind(outliersremove,outliersmerge)
alloutliers<-as.vector(alloutliers[,1])

shape2016b<-shape2016[!shape2016$CTNAME %in% alloutliers,]
shape2016b<-spChFIDs(shape2016b, as.character(shape2016b$CTNAME))
newshape<-shape2016b[,(2)]

for (i in 1:nrow(merge)) {
test1<-shape2016[shape2016$CTNAME==as.numeric(merge[i,1]),]
test2<-shape2016[shape2016$CTNAME==as.numeric(merge[i,2]),]
test1$CTNAME<-as.numeric(merge[i,2])
test3<-rbind(test1,test2)
test4<-aggregate(test3, by="CTNAME")
newshape<-rbind(newshape, test4)
rm(test1,test2,test3,test4)
}

rm(shape2016,shape2016b,outliersremove,outliersmerge,alloutliers,i,merge)

#function to iterate the shape merging

dir<-"R:/working/preparing data/Merge"
mergeShape<-function(filename) {
  merge<-read.csv(paste0(dir, "/",filename ,".csv"), header=T)
  outliersremove<-dplyr::select(merge,1)
  outliersmerge<-dplyr::select(merge,2)
  names(outliersremove)<-names(outliersmerge)
  alloutliers<-rbind(outliersremove,outliersmerge)
  alloutliers<-as.vector(alloutliers[,1])
  
  shape2<-newshape[!newshape$CTNAME %in% alloutliers,]
  shape2<-spChFIDs(shape2, as.character(shape2$CTNAME))
  newshape2<-shape2[,(1)]
  
  for (i in 1:nrow(merge)) {
  test1<-newshape[newshape$CTNAME==as.numeric(merge[i,1]),]
  test2<-newshape[newshape$CTNAME==as.numeric(merge[i,2]),]
  test1$CTNAME<-as.numeric(merge[i,2])
  test3<-rbind(test1,test2)
  test4<-aggregate(test3, by="CTNAME")
  newshape2<-rbind(newshape2, test4)
  rm(test1,test2,test3,test4)
  }
  return(newshape<-newshape2)
}

newshape<-mergeShape("merge_2")
newshape<-mergeShape("merge_3")
newshape<-mergeShape("merge_4")
newshape<-mergeShape("merge_5")
newshape<-mergeShape("merge_6")
newshape<-mergeShape("merge_7")
newshape<-mergeShape("merge_8")

newshape<-subset(newshape, !CTNAME==270)
shape<-newshape[order(newshape@data[["CTNAME"]]),]


#Save CTIDs to identify shapes later
CT<-unique(shape@data[["CTNAME"]])
numbers<-0:(length(shape)-1)
CT<-data.frame(numbers,CT)
names(CT)<-c("area","CT")
save(CT, file="R:/working/preparing data/CT368.RDATA")

shape$CTNAME<-1:nrow(shape) #changing so that ct names are sequential from 1
subtract = 1
shape$CTNAME<-shape$CTNAME - subtract

#change shape id to reflect rowname
for (i in 1:nrow(shape))
shape@polygons[[i]]@ID<-as.character(i-1)

plot(shape)
save(shape, file="R:/working/modeling/shape_files/shape-mapping368CTs.RData")

rm(newshape, merge, outliersmerge, outliersremove, shape2016, shape2016b, alloutliers, i, subtract,newshape2,shape2)

```


