---
title: "adjustSubsetDeathCTs"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                      
             (B) Recode death files of interest and save file                                                   

## Load libraries, import data, and define attributes

```{r}
libs<-c("data.table","dplyr", "tidyr")
lapply(libs, require, character.only=TRUE)
load("R:/working/preparing data/DeathData/causes/subset/deaths_cwapplied_subset.RData")
merge<-read.csv("R:/working/preparing data/Merge/merge_all12.csv", header=T)

library(car)
for (i in 1:nrow(merge)){
  deaths$CT[which(deaths$CT==as.numeric(merge[i,1]))]<-as.numeric(merge[i,2])
}
detach("package:car", unload=TRUE)
library(dplyr)
```


## Recode death files of interest and save file

```{r}

newDeaths <- deaths %>%
  group_by(CT,year,sex,age) %>%
  summarise(death=sum(death, na.rm=TRUE)) %>%
  mutate(death=floor(death))
newDeaths<-subset(newDeaths,CT!=270)

deaths<-newDeaths[order(newDeaths$CT),]
deaths$CT<-as.numeric(as.factor(deaths$CT))
deaths$CT<-as.numeric(deaths$CT)
deaths$CT<-(deaths$CT-1)

deaths$sex<-as.integer(deaths$sex)
deaths<-na.omit(deaths)
deaths<-data.table::setDT(deaths)

length(unique(deaths$CT))

save(deaths, file="R:/working/modeling/deaths/deaths_368CTs_subset.RData")

```

