---
title: "categoriseSubsetNeoDeaths"
output: html_document
---

  Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                       
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load the relevant libraries and import data                                                  
             (B) Categorise age groups
             (C) Keep final columns 
             (D) Derive causes using ICD codes for NEOPLASMS
             (E) Count number of deaths and save file
             
## Load the relevant libraries and import data

```{r}
library("dplyr")
library("tidyr")
library("data.table")

setwd("R:/working/preparing data/DeathData")
allDeaths<-read.csv(file="final_mv_allyears.csv", header=TRUE, check.names = FALSE)
```


## Categorise age groups

```{r}
#Categorise age groups and rename column names
agebreaks <-c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,113)
agelabels <-c("0","5","10","15","20","25","30","35","40","45","50","55","60","65","70","75","80","85")

deaths<-setDT(allDeaths) [, agegroups := cut(AGE, breaks = agebreaks, right= FALSE, labels = agelabels)]

```

## Keep final columns and change variable types

```{r}

colnames(deaths)[colnames(deaths)=="DEATHDTCCYY"]<-"year"
#colnames(deaths)[colnames(deaths)=="SEX"]<-"sex"
colnames(deaths)[colnames(deaths)=="agegroups"]<-"age"
colnames(deaths)[colnames(deaths)=="CAUSE"]<-"cause"
colnames(deaths)[colnames(deaths)=="ICD_VERSN"]<-"icdversn"

deaths <- deaths %>% select(year,CT,SEX,age,cause,icdversn) %>% mutate(sex=case_when(
                             SEX =="F" ~ "2",
                             SEX == "M"~ "1",
                             SEX == "U" ~ "3",
                             TRUE ~ "other"))

deaths$sex<-as.numeric(as.character(deaths$sex))
deaths$age<-as.numeric(as.character(deaths$age))

deaths <-select(deaths,year,CT,sex,age,cause,icdversn)

rm(allDeaths, agebreaks, agelabels)

```


#Derive causes using ICD codes for NEOPLASMS - ISCHEMIC and STROKE in another folder for categorisation

```{r}
deaths$cause<-trimws(deaths$cause)

deaths<-deaths %>% mutate(causeid=
                                 case_when(cause =="C349" | cause=="1629" ~ "lung",
                                           cause =="C61" | cause=="185" ~ "prostate",
                                           cause =="C189" | cause=="1539" ~ "colon",
                                           cause =="C259" | cause=="1579" ~ "pancreas",
                                           cause =="C159" | cause=="1509" ~ "esophagus",
                                           cause =="C509" | cause=="1749" ~ "breast",
                                           cause =="C56" | cause=="1830" ~ "ovary"                                           ))

deaths_lung<- deaths %>% filter(causeid=="lung")
deaths_prostate<- deaths %>% filter(causeid=="prostate")
deaths_colon<- deaths %>% filter(causeid=="colon")
deaths_pancreas<- deaths %>% filter(causeid=="pancreas")
deaths_esophagus<- deaths %>% filter(causeid=="esophagus")
deaths_breast<- deaths %>% filter(causeid=="breast")
deaths_ovary<- deaths %>% filter(causeid=="ovary")


deaths<-deaths_ovary

```


## Count number of deaths and save file

```{r}
#Count number of deaths
mortcounts<-count(deaths, year, CT, sex, age)
colnames(mortcounts)[colnames(mortcounts)=="n"]<-"death"

#sort data by sequentially of CT
deaths<-mortcounts[order(mortcounts$CT),]


save(deaths, file="R:/working/preparing data/DeathData/causes/subset/deaths_subset.rdata")

```

