---
title: "MergeCTs"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                                         
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium                                                                                                          
   Paper published from this work: 
   Yu J, Dwyer-Lindgren L, Bennett J, Ezzati M, Gustafson P, Tran M, Brauer M. A spatiotemporal analysis of inequalities in life expectancy and 20 causes of mortality in sub-neighbourhoods of Metro Vancouver, British Columbia, Canada, 1990–2016. Health & place. 2021 Nov 1;72:102692. Available from: https://doi.org/10.1016/j.healthplace.2021.102692                                                    
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries and import data                                                 
             (B) Find which CTs have deaths less than 1 for each CT-year      
             (C) Merge CTs that have less than one death and greater than 0                      
             (D) Find out which CTs have pop less than 150 per census year 
             (E) Combine CT lists and ensure no duplicates 
             (F) Merge and separate CTs 
             (G) Find CTs with mx in the last age group that is less than 0.06 to merge
             (H) Combine CT lists and ensure no duplicates
             
#Load libraries and import data

```{r}
libs<-c("car","data.table", "dplyr")
lapply(libs, require, character.only=TRUE)
load("R:/working/preparing data/PopulationData/pop_interpolated.RDATA")
load("R:/working/modeling/deaths/old/deaths_cwapplied.RData")
deaths<-subset(deaths, CT!=270)



```

##Find which CTs have deaths less than 1 for each CT-year

```{r}
deaths<-subset(deaths, CT!=270)

deathInfo <- deaths %>%
  group_by(year,CT) %>%
  summarise(deaths=sum(death))

merge<-subset(deathInfo, deaths<=1 & year<2001)
merge<- merge %>%
  group_by(CT)%>%
  summarise()
names(merge)<-"ctuid_t"

CTs<-merge

crosswalk91<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk91.csv", header=TRUE)

crosswalk91<-select(crosswalk91,2,3)
crosswalk91b<-select(crosswalk91, 2) #isolate target CTs to know which 2016 cts to merge back to
merge2<-setdiff(crosswalk91b,merge)
crosswalk91b<-left_join(merge2,crosswalk91)
names(crosswalk91b)[1]<-"ctuid_t2"


merge<-merge(merge,crosswalk91,by="ctuid_t")
merge<-left_join(merge,crosswalk91b,by="ctuid_s")
names(merge)[3]<-"CT"
merge<-within(merge, CT[is.na(CT)]<-ctuid_t[is.na(CT)])

deathInfo<-deathInfo %>%
  group_by(CT) %>%
  summarise(deaths=sum(deaths))
deathInfo2<-select(deathInfo,"CT","deaths")
merge<-left_join(merge,deathInfo2, by="CT")

merge$choice<-0
CTs$choice<-0

#Make a decision of CT based on the one with smallest population
for (i in 1:nrow(CTs)){
  x<-which(merge$ctuid_t==CTs$ctuid_t[i])
  y<-min(merge$deaths[x])
  merge$choice[x]<-merge$CT[which(merge$deaths==y)]
  CTs$choice[i]<-merge$CT[which(merge$deaths==y)]
}

mergeDeaths<-merge
merge2<-CTs

names(merge2)<-c("outliersremove","outliersmerge")

write.csv(merge2, file="R:/working/preparing data/merge/mergeDeathList.csv", row.names = FALSE)

```

##Merge CTs that have less than one death and greater than 0 - merge with neighbour CT with more deaths 

```{r}
load("R:/working/modeling/deaths/old/deaths_cwapplied.RData")
deaths<-subset(deaths, CT!=270)

deathInfo <- deaths %>%
  group_by(year,CT) %>%
  summarise(deaths=sum(death))

merge<-subset(deathInfo, deaths<1)
merge<- merge %>%
  group_by(CT)%>%
  summarise()
names(merge)<-"outliersremove"

CTs<-unique(merge)


#Load shapefile and find polygon association
library(sf)
load("R:/working/modeling/shape_files/shape_478CTs.RData")
library("spdep")
#find neighbours
shape<-st_read("R:/working/preparing data/ShapeFiles/2016MetroVanCTs.shp")
CT<-data.frame(CT=shape$CTNAME,area=1:478)
CT$CT<-as.numeric(as.character(CT$CT))

#look for CTs that need another CT to merge with because it's original CT
merge<-merge %>%
  rename(CT=outliersremove) %>%
  left_join(CT,by="CT") %>%
  select("area")

#Find neighbours
NB <- poly2nb(shape, queen=FALSE, row.names=shape$CTNAME)

subsetList<-function(NB, elementNames) {
  lapply(elementNames, FUN=function(x) CT$CT[NB[[x]]])
}

#Create list of CTs that are neighbours to ones of interest
for (i in merge){
NB2<-subsetList(NB, c(i))
}

load("R:/working/modeling/deaths/old/deaths_cwapplied.RData")
#Create death file of interest
deathInfo <- deaths %>%
  group_by(CT,sex) %>%
  summarise(deaths=sum(death)) %>%
  filter(sex==2) %>%
  rename(x=CT)

#Link neighbourhood file with death data to identify CT neighbours with more deaths
NB3<-lapply(NB2, function(x) merge(x, deathInfo, by="x"))
new.names<-c("CT","sex", "deaths")
NB4<-lapply(NB3, setNames, new.names)  
NB5<-lapply(NB4, function(x) subset(x, deaths==min(deaths)))
library(purrr)
NB6<-map(NB5, ~(.x %>% select(CT)))
deathCTchoice<-do.call(rbind.data.frame, NB6)

#link CTs to merge from neighbour list
merge2<-CTs %>%
  cbind(deathCTchoice) %>%
  rename(outliersmerge=CT)

write.csv(merge2, file="R:/working/preparing data/merge/mergeDeathList.csv", row.names = FALSE)


```


##Find out which CTs have pop less than 150 per census year and merge with original CT

```{r}

load("R:/working/preparing data/PopulationData/pop_interpolated.RDATA")
popInfo <- pop %>%
  group_by(year,CT,sex) %>%
  summarise(pop=sum(pop))

#Identify CTs with pop less than 150
merge<-subset(popInfo, pop<=300)
merge<- merge %>%
  group_by(CT)%>%
  summarise()
names(merge)<-"ctuid_t"

CTs<-merge

#load crosswalk to isolate CTs to know which 2016 CTs to merge back to (ideally original CT)
crosswalk91<-read.csv(file="R:/working/preparing data/prepareCrosswalk/prepared/crosswalk91.csv", header=TRUE)
crosswalk91<-crosswalk91 %>%
  select(2,3)
crosswalk91b<-select(crosswalk91, 2) #isolate target CTs to know which 2016 cts to merge back to
merge2<-setdiff(crosswalk91b,merge)
crosswalk91b<-left_join(merge2,crosswalk91)
names(crosswalk91b)[1]<-"ctuid_t2"

#Create CT column identify which CT to merge to
merge<-merge(merge,crosswalk91,by="ctuid_t")
merge<-left_join(merge,crosswalk91b,by="ctuid_s")
names(merge)[3]<-"CT"
merge<-within(merge, CT[is.na(CT)]<-ctuid_t[is.na(CT)]) #If no CT to merge to because it's the original CT, then replace to original CT

#Merge with 1991 population to know which CTs to merge with (try to merge with CT neighbour with less population)
popInfo<-subset(popInfo, year==1991)
popInfo2<-select(popInfo,"CT","pop")
merge<-left_join(merge,popInfo2, by="CT")
merge<-select(merge,-"year")

#Create choice column for decision making
merge$choice<-0
CTs$choice<-0

#Make a decision of CT based on the one with smallest population
for (i in 1:nrow(CTs)){
  x<-which(merge$ctuid_t==CTs$ctuid_t[i])
  y<-min(merge$pop[x])
  merge$choice[x]<-merge$CT[which(merge$pop==y)]
  CTs$choice[i]<-merge$CT[which(merge$pop==y)]
}

merge2<-merge
merge<-CTs

names(merge)<-c("outliersremove","outliersmerge")

write.csv(merge, file="R:/working/preparing data/Merge/merge.csv", row.names = FALSE) 

```

#Combine CT lists and ensure no duplicates 

```{r}

merge<-read.csv("R:/working/preparing data/Merge/merge.csv", header=T)
merge2<-read.csv("R:/working/preparing data/Merge/mergeDeathList.csv", header=T)

uniqueCTs<-function(df,df2) {
  df1.1<-df %>%
    select(1)
  df2.1<-df2 %>%
    select(1) 
  same.1<-intersect(df1.1,df2.1) #same outliersmerge
  
  df1.2<-df %>%
    select(2)
  df2.2<-df2 %>%
    select(2)
  names(df1.1)<-"outliersmerge"
  same.2<-intersect(df1.1,df2.2) #outliersmerge in df2 that is similar to df1

  same.3<-intersect(df1.1,df1.2) #outliersmerge in df1 that is similar to outliersremove in df2
  
  names(df2.1)<-"outliersmerge"
  same.4<-intersect(df2.1, df2.2) #outliersmerge in df2 that is similar to outliersremove in df2
  
  same.5<-intersect(df2.1, df1.2) #outliersmerge in df2 that is similar to outliersremove in df2

  return(list(same.1,same.2,same.3,same.4,same.5)) 
}

checkSame<-uniqueCTs(merge,merge2)

#remove outliersremove that are same for pop and deaths - choose pop ones
merge3<-merge2 %>%
  filter(!outliersremove %in% checkSame[[1]][["outliersremove"]])

checkSame<-uniqueCTs(merge,merge3)

#look for CTs that need to swap in second df
merge4<-merge3 %>%
  filter(outliersmerge %in% checkSame[[2]][["outliersmerge"]])
names(merge4)<-c("outliersmerge","outliersremove")
merge4<-select(merge4,2,1)

#remove CTs that need to swap and then add new df created that's swapped
merge3<-merge3 %>%
  filter(!outliersmerge %in% checkSame[[2]][["outliersmerge"]])
merge3<-rbind(merge3,merge4)
rm(merge4)

checkSame<-uniqueCTs(merge,merge3)

#remove outliersremove that are same for pop and deaths - choose pop ones
merge3<-merge3 %>%
  filter(!outliersremove %in% checkSame[[1]][["outliersremove"]])

checkSame<-uniqueCTs(merge,merge3)


#manually remove from df1 first row that is merged with 251.02 and impute 251.01

merge<-merge %>%
  filter(!outliersremove %in% checkSame[[3]][["outliersmerge"]][1])

for (i in checkSame[[3]][["outliersmerge"]][2]) {
  merge$outliersmerge[which(merge$outliersremove==i)]<-checkSame[[3]][["outliersmerge"]][1]
}

checkSame<-uniqueCTs(merge,merge3)


#Load shapefile and find polygon association
library(sf)
load("R:/working/modeling/shape_files/shape_478CTs.RData")
library("spdep")
#find neighbours
shape<-st_read("R:/working/preparing data/ShapeFiles/2016MetroVanCTs.shp")
CT<-data.frame(CT=shape$CTNAME,area=1:478)
CT$CT<-as.numeric(as.character(CT$CT))

#look for CTs that need another CT to merge with because it's original CT
merge4<-merge3 %>%
  filter(outliersmerge %in% checkSame[[4]][["outliersmerge"]]) %>%
  select(1) %>%
  rename(CT=outliersremove) %>%
  left_join(CT,by="CT") %>%
  select("area")

#Find neighbours
NB <- poly2nb(shape, queen=TRUE, row.names=shape$CTNAME)

subsetList<-function(NB, elementNames) {
  lapply(elementNames, FUN=function(x) CT$CT[NB[[x]]])
}

#Create list of CTs that are neighbours to ones of interest
for (i in merge4){
NB2<-subsetList(NB, c(i))
}

load("R:/working/modeling/deaths/old/deaths_cwapplied.RData")
#Create death file of interest
deathInfo <- deaths %>%
  filter(year<2009) %>%
  group_by(CT,sex) %>%
  summarise(deaths=sum(death)) %>%
  filter(sex==2) %>%
  rename(x=CT)

#Link neighbourhood file with death data to identify CT neighbours with more deaths
NB3<-lapply(NB2, function(x) merge(x, deathInfo, by="x"))
new.names<-c("CT","sex", "deaths")
NB4<-lapply(NB3, setNames, new.names)  
NB5<-lapply(NB4, function(x) subset(x, deaths==min(deaths)))
library(purrr)
NB6<-map(NB5, ~(.x %>% select(CT)))
deathCTchoice<-do.call(rbind.data.frame, NB6)

#identify CTs to merge from neighbour list
merge4<-merge3 %>%
  filter(outliersmerge %in% checkSame[[4]][["outliersmerge"]]) %>%
  select(1) %>%
  cbind(deathCTchoice) %>%
  rename(outliersmerge=CT)

merge3<-merge3 %>%
  filter(!outliersremove %in% checkSame[[4]][["outliersmerge"]]) %>%
  rbind(merge4)

checkSame<-uniqueCTs(merge,merge3) 

#look for CTs that need to swap in second df
merge4<-merge3 %>%
  filter(outliersremove %in% checkSame[[5]][["outliersmerge"]])
names(merge4)<-c("outliersmerge","outliersremove")
merge4<-select(merge4,2,1)

#remove CTs that need to swap and then add new df created that's swapped
merge3<-merge3 %>%
  filter(!outliersremove %in% checkSame[[5]][["outliersmerge"]])
merge3<-rbind(merge3,merge4)
rm(merge4)

checkSame<-uniqueCTs(merge,merge3)

#look for CTs that need to swap in second df
merge4<-merge3 %>%
  filter(outliersremove %in% checkSame[[4]][["outliersmerge"]])
names(merge4)<-c("outliersmerge","outliersremove")
merge4<-select(merge4,2,1)

#remove CTs that need to swap and then add new df created that's swapped
merge3<-merge3 %>%
  filter(!outliersremove %in% checkSame[[4]][["outliersmerge"]][1])
merge3<-rbind(merge3,merge4)
rm(merge4)

#Remove one row that are the same but just swapped
merge3<-merge3 %>%
  filter(!outliersmerge %in% checkSame[[2]][["outliersmerge"]])

checkSame<-uniqueCTs(merge,merge3) #reveals same problem, so we need to impute the outliersmerge numbers to outliersremove 


merge_all<-rbind(merge,merge3)
merge_all<-merge_all[order(merge_all$outliersremove),]

for (i in checkSame[[5]][["outliersmerge"]]) {
  merge_all$outliersmerge[which(merge_all$outliersmerge==i)]<-merge_all$outliersmerge[which(merge_all$outliersremove==i)]
}



checkSame<-intersect(merge_all$outliersremove,merge_all$outliersmerge) #last check
length(unique(merge_all$outliersremove))#this should equal to same length of merge_all

notUnique<-merge_all$outliersremove[which(duplicated(merge_all$outliersremove)==TRUE)]
notUniqueRows<-which(duplicated(merge_all$outliersremove)==TRUE)

merge_all<-merge_all %>%
  filter(!row_number() %in% notUniqueRows)

write.csv(merge_all, file="R:/working/preparing data/Merge/merge_all3.csv", row.names = F)

```


#Merge and separate CTs

```{r}
library(tidyverse)
merge_all<-read.csv("R:/working/preparing data/Merge/merge_all3.csv", header=T)
merge.all<-aggregate(outliersremove ~ outliersmerge, merge_all, paste, collapse = "; ")

merge.split<-separate(merge.all, col=outliersremove, into=c("outliers1", "outliers2", "outliers3", "outliers4"), sep=";")

merge.split<-separate(merge.all, col=outliersremove, into=c("outliers1", "outliers2", "outliers3", "outliers4","outliers5","outliers6","outliers7","outliers8"), sep=";")
library(dplyr)
merge.1<-merge.split %>%
  select(2,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers1)

merge.2<-merge.split %>%
  select(3,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers2)

merge.3<-merge.split %>%
  select(4,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers3)

merge.4<-merge.split %>%
  select(5,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers4)

merge.5<-merge.split %>%
  select(6,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers5)

merge.6<-merge.split %>%
  select(7,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers6)

merge.7<-merge.split %>%
  select(8,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers7)

merge.8<-merge.split %>%
  select(9,1) %>%
  na.omit() %>%
  rename(outliersremove=outliers8)

  
write.csv(merge.1, file="R:/working/preparing data/Merge/merge_1.csv", row.names=F) 
write.csv(merge.2, file="R:/working/preparing data/Merge/merge_2.csv", row.names=F) 
write.csv(merge.3, file="R:/working/preparing data/Merge/merge_3.csv", row.names=F) 
write.csv(merge.4, file="R:/working/preparing data/Merge/merge_4.csv", row.names=F)
write.csv(merge.5, file="R:/working/preparing data/Merge/merge_5.csv", row.names=F)
write.csv(merge.6, file="R:/working/preparing data/Merge/merge_6.csv", row.names=F)
write.csv(merge.7, file="R:/working/preparing data/Merge/merge_7.csv", row.names=F)
write.csv(merge.8, file="R:/working/preparing data/Merge/merge_8.csv", row.names=F)


```


#Find CTs with mx in the last age group that is less than 0.06 to merge

```{r}
#Need LE estimates from running last SAE models - 447 CTs

load("R:/working/modeling/temp_dir/lt_est_1991.rdata")
est1990<-est

mxCheckF<- est1990 %>%
  filter(age==85 & sex==2) %>%
  select(area,mx_mean) %>%
  group_by(area) %>%
  summarise(mx_mean=mean(mx_mean)) 

mxCheckM<- est1990 %>%
  filter(age==85 & sex==1) %>%
  select(area,mx_mean) %>%
  group_by(area) %>%
  summarise(mx_mean=mean(mx_mean)) 

mergeF<-mxCheckF %>%
  filter(mx_mean<0.06) %>%
  select(area)

mergeM<-mxCheckM %>%
  filter(mx_mean<0.06) %>%
  select(area)

merge1990<-mergeF %>%
  rbind(mergeM) %>%
  unique()

load("R:/working/modeling/temp_dir/lt_est_2016.rdata")
est2016<-est

mxCheckF<- est2016 %>%
  filter(age==85 & sex==2) %>%
  select(area,mx_mean) %>%
  group_by(area) %>%
  summarise(mx_mean=mean(mx_mean)) 

mxCheckM<- est2016 %>%
  filter(age==85 & sex==1) %>%
  select(area,mx_mean) %>%
  group_by(area) %>%
  summarise(mx_mean=mean(mx_mean)) 

mergeF<-mxCheckF %>%
  filter(mx_mean<0.06) %>%
  select(area)

mergeM<-mxCheckM %>%
  filter(mx_mean<0.06) %>%
  select(area)

merge2016<-mergeF %>%
  rbind(mergeM) %>%
  unique()

#Load CTs to merge with areas of interest to merge

load("R:/working/preparing data/CT369.RDATA")
names(CT)<-c("area", "CT")

mergeBoth<-merge1990 %>%
  rbind(merge2016) %>%
  unique() %>%
  left_join(CT, by="area") %>%
  select(CT) %>%
  arrange(CT)

#load shapefile to find neighbours to merge with
load("R:/working/modeling/shape_files/shape-mapping369CTs.RData")
library("sf")
library("spdep")
#find neighbours
load("R:/working/preparing data/CT369.RDATA")
names(CT)<-c("area", "CT")
CT<-CT %>%
  mutate(area=area+1)

#look for CTs that need another CT to merge with because it's original CT
merge<-mergeBoth %>%
  left_join(CT,by="CT") %>%
  select("area")

#Find neighbours
NB <- poly2nb(shape, queen=FALSE, row.names=shape$CTNAME)

subsetList<-function(NB, elementNames) {
  lapply(elementNames, FUN=function(x) CT$CT[NB[[x]]])
}

#Create list of CTs that are neighbours to ones of interest
for (i in merge){
NB2<-subsetList(NB, c(i))
}

#Find CTs with low or high mortality rates in the last age group to merge with
load("R:/working/modeling/temp_dir/lt_est_2016.rdata")
est1990<-est

load("R:/working/preparing data/CT373.RDATA")
names(CT)<-c("area", "CT")
mxCheckF<- est1990 %>%
  filter(age==85 & sex==2) %>%
  left_join(CT) %>%
  select(CT,mx_mean) %>%
  group_by(CT) %>%
  summarise(mx_mean=mean(mx_mean)) %>%
  rename(x=CT)

#Link neighbourhood file with death data to identify CT neighbours with more deaths
NB3<-lapply(NB2, function(x) merge(x, mxCheckF, by="x"))
new.names<-c("CT","mx_mean")
NB4<-lapply(NB3, setNames, new.names)  
  
NB4[[6]][["mx_mean"]][which(NB4[[6]][["CT"]]==150)]<-0
NB4[[9]][["mx_mean"]][which(NB4[[9]][["CT"]]==192)]<-0
NB4[[17]][["mx_mean"]][which(NB4[[17]][["CT"]]==501.01)]<-0


NB5<-lapply(NB4, function(x) subset(x, mx_mean==max(mx_mean))) #max mx
library(purrr)
NB6<-map(NB5, ~(.x %>% select(CT)))
MXCTchoice<-do.call(rbind.data.frame, NB6)

#link CTs to merge from neighbour list
merge6<-mergeBoth %>%
  rename(outliersremove=CT) %>%
  cbind(MXCTchoice) %>%
  rename(outliersmerge=CT)

#manually take out airport CT (150) from outliersmerge
merge6$outliersmerge[which(merge6$outliersremove==149.07)]<-149.08

#remove link to 501.01

write.csv(merge6, file="R:/working/preparing data/Merge/lowMXCts_3.csv", row.names=F)
```

#Combine CT lists and ensure no duplicates 

```{r}

merge<-read.csv("R:/working/preparing data/Merge/merge_all7.csv", header=T)
merge2<-read.csv("R:/working/preparing data/Merge/lowMXCts_3.csv", header=T)

uniqueCTs<-function(df,df2) {
  df1.1<-df %>%
    select(1)
  df2.1<-df2 %>%
    select(1) 
  same.1<-intersect(df1.1,df2.1) #same outliersmerge
  
  df1.2<-df %>%
    select(2)
  df2.2<-df2 %>%
    select(2)
  names(df1.1)<-"outliersmerge"
  same.2<-intersect(df1.1,df2.2) #outliersmerge in df2 that is similar to df1

  same.3<-intersect(df1.1,df1.2) #outliersmerge in df1 that is similar to outliersremove in df2
  
  names(df2.1)<-"outliersmerge"
  same.4<-intersect(df2.1, df2.2) #outliersmerge in df2 that is similar to outliersremove in df2
  
  same.5<-intersect(df2.1, df1.2) #outliersremove in df2 that is similar to outliersmerge in df1

  return(list(same.1,same.2,same.3,same.4,same.5)) 
}

checkSame<-uniqueCTs(merge,merge2)

#look for CTs that need to swap in second df
merge4<-merge2 %>%
  filter(outliersremove %in% checkSame[[4]][["outliersmerge"]])
names(merge4)<-c("outliersmerge","outliersremove")
merge4<-select(merge4,2,1)

#remove CTs that need to swap and then add new df created that's swapped
merge3<-merge2 %>%
  filter(!outliersremove %in% checkSame[[4]][["outliersmerge"]])
merge3<-rbind(merge3,merge4)
rm(merge4)

checkSame<-uniqueCTs(merge,merge3)

#change first merge list in outliersmerge to match second merge list

merge3<-merge2

for (i in checkSame[[5]][["outliersmerge"]]) {
  merge$outliersmerge[which(merge$outliersmerge==i)]<-merge3$outliersmerge[which(merge3$outliersremove==i)]
}

checkSame<-uniqueCTs(merge,merge3)


merge_all<-rbind(merge,merge3)
merge_all<-merge_all[order(merge_all$outliersremove),]

checkSame<-intersect(merge_all$outliersremove,merge_all$outliersmerge) #last check
length(unique(merge_all$outliersremove))#this should equal to same length of merge_all

write.csv(merge_all, file="R:/working/preparing data/Merge/merge_all10.csv", row.names = F)

```
